---
title: "1: Dowload RAW data"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> Download raw data using GEOquery
</div>

### Background 

We are going to download raw data and meta information from [GEO](https://www.ncbi.nlm.nih.gov/geo/) using `GEOquery` package.


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>GSE list:</strong> GSE IDs from the experiments
</div>


## Setup

For running the script, type:

```{bash, eval = F}
ssh -X aaltamirano@dna.lavis.unam.mx
qrsh
cd /mnt/Genoma/amedina/aaltamirano/COPD/Meta-analysis_COPD/output_data/.out
 module load r/4.0.1 

nohup R -e "rmarkdown::render(here::here('vignettes/1-download_raw-data.RMD'))" > 1-download_raw-data.out &
```



The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.

```{r setup}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("R/setup.R"))
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(GEOquery)
```



## Download raw data

We need the GSE IDs to download raw files, the IDs are stored in a separate file named `GSE_IDs.txt`, we used `getGEOSuppFiles()` for download raw data. We do it using GSE ids from tissue:

```{r , class.source="bg-info",class.output="bg-info" }
tissue <- scan(DATA("tissue_GSE_IDs.txt"),
               character(), 
               sep = ",")

sapply(tissue,getGEOSuppFiles,baseDir = here::here("GEOquery"))
```

## Download GEOquery data

We download GEOquery data for the experiments 

```{r}
geo <- sapply(tissue,getGEO,destdir = here::here("GEOquery"))
```

Some GSE ids have more than one platform and samples that we will not use, so we delete them or do some modifications. 

```{r}
sapply(geo,length)
# we want the experiment with 64 samples, so we reverse the element
geo[[1]] <- rev(geo[[1]])

saveRDS(geo,OUTPUT(c(TODAY,"_GEOquery-download.RDS")))
```


```{bash, eval =F}
cd GEOquery/GSE27597/
rm GSM6844*
rm GSM6845*

# Divide samples from differnt platforms
find . -size -3000k -type f | wc -l
find . -size -3000k -type f -exec mv "{}" 153_samples/ \;

```


## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Data 1:</strong> RAW data in  `/GEOquery`  
  <strong>RDS:</strong> Data from GEOquery `OUTPUT(c(TODAY,"_GEOquery-download.RDS"))` 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





