---
title: "3-curating-data.RMD"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To manually curate sample information and to join meta data with normalized data
</div>

We previously had run the following scripts:

- 0: Data selection `/0-Data-selection.RMD`

- 1: Dowload RAW data `/1-download_raw-data.RMD`

- 2: Normalizing data `/2-normalizing-data.RMD`

## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Data 1:</strong> Data information from geo 2020-09-12_GEOquery-download.RDS     
  <strong>Data 2:</strong> Matrix of normalized experiments 2020-09-24_normData.xz  
</div>



## Setup

For running the script, type:

```{bash, eval = F}
ssh -X aaltamirano@dna.lavis.unam.mx
qrsh
cd /mnt/Genoma/amedina/aaltamirano/COPD/Meta-analysis_COPD/output_data/.out
 module load r/4.0.1 

## OR
ssh ana@10.200.0.42
cd /home/ana/DataDNA/R-projects/Meta-analysis_COPD/output_data/.out

nohup R -e "rmarkdown::render(here::here('vignettes/3-curatig-data.RMD'))" > 3-curatig-data.RMD.out &
```

The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.

```{r setup}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("Tissue/R/setup.R"))
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(GEOquery)
library(mgsub)
library(annotate)
library(hugene10sttranscriptcluster.db)
library(EnsDb.Hsapiens.v79)
library(tidyverse)
library(gsubfn)
```

### Function

Function to obtain sample information after curated and stored in a tibble

```{r}

obtainInfo <- function(df){
  df <- df %>% add_row(
          "GSE" = names(norm)[i],
          "CONTROL" = table(p$DISEASE)[1],
          "COPD" = table(p$DISEASE)[2],
          "OTHER" = sum(table(p$DISEASE)[-c(1,2)]),
          "COUNTRY" = unique(p$contact_country),
          "SUBMISSION_DATE" = paste(unique(p$submission_date),collapse = ","),
          "PLATFORM" = unique(p$platform_id),
          "NUM_GENES" = length(unique(f$GENE.SYMBOL)),
          "H_AGE_MEAN" = mean(as.numeric(p$AGE[which(p$DISEASE=="CONTROL")]),na.rm = T),
          "H_AGE_MEDIAN" = median(as.numeric(p$AGE[which(p$DISEASE=="CONTROL")]),na.rm = T),
          "H_AGE_RANGE" = paste(as.character(range(as.numeric(p$AGE[which(p$DISEASE=="CONTROL")]),na.rm = T)),collapse= "-"),
          "H_SEX_F" = table(p$SEX[which(p$DISEASE=="CONTROL")])[1],
          "H_SEX_M" = table(p$SEX[which(p$DISEASE=="CONTROL")])[2],
          "H_SMOKING_STATUS" = toString(c(names(table(p$SMOKING_STATUS[which(p$DISEASE=="CONTROL")])),table(p$SMOKING_STATUS[which(p$DISEASE=="CONTROL")]))),
          "H_PACKYEAR" = median(na.omit(as.numeric(p$PACKYEAR[which(p$DISEASE=="CONTROL")]))),
          "COPD_AGE_MEAN" = mean(as.numeric(p$AGE[which(p$DISEASE=="COPD")]),na.rm = T),
          "COPD_AGE_MEDIAN" = median(as.numeric(p$AGE[which(p$DISEASE=="COPD")]),na.rm = T),
          "COPD_AGE_RANGE" = paste(as.character(range(as.numeric(p$AGE[which(p$DISEASE=="COPD")]),na.rm = T)),collapse= "-"),
          "COPD_SEX_F" = table(p$SEX[which(p$DISEASE=="COPD")])[1],
          "COPD_SEX_M" = table(p$SEX[which(p$DISEASE=="COPD")])[2],
          "COPD_SMOKING_STATUS" = toString(c(names(table(p$SMOKING_STATUS[which(p$DISEASE=="COPD")])),table(p$SMOKING_STATUS[which(p$DISEASE=="COPD")]))),
          "COPD_PACKYEAR" = median(na.omit(as.numeric(p$PACKYEAR[which(p$DISEASE=="COPD")])))
  )
}

```




## Curation

### STEP 1: Import data

We read GEOquery data for all the experiments, it has the 

```{r}
geo <- readRDS(OUTPUT("2020-09-12_GEOquery-download.RDS"))
norm <- readRDS(OUTPUT("2020-09-24_normData.xz"))

names(geo)
names(norm)
```


### STEP 2: Curate information


**NEW COLUMN and CONTROLED VOCABULARY!!**

We will add a new column and use controlled vocabulary, so we will change the 
description that each authors used to describe COPD or Control samples.

In addition, we will create a new table summarizing data from all experiments, 
it will contain number of COPD samples, controls, experiment publication date, etc. 

```{r}
######## SUMMARY TABLE
## Summary information about the experiment
df <- tibble("GSE"= NA,
             "CONTROL" = NA,
             "COPD" = NA,
             "OTHER" = NA,
             "COUNTRY" = NA,
             "SUBMISSION_DATE" = NA,
             "PLATFORM" = NA,
             "NUM_GENES" = NA,
             "H_AGE_MEAN" = NA,
             "H_AGE_MEDIAN" = NA,
             "H_AGE_RANGE" = NA,
             "H_SEX_F" = NA,
             "H_SEX_M" = NA,
             "H_SMOKING_STATUS" = NA,
             "H_PACKYEAR" = NA,
             "COPD_AGE_MEAN" = NA,
             "COPD_AGE_MEDIAN" = NA,
             "COPD_AGE_RANGE" = NA,
             "COPD_SEX_F" = NA,
             "COPD_SEX_M" = NA,
             "COPD_SMOKING_STATUS" = NA,
             "COPD_PACKYEAR" = NA
                 )

```


## `r names(geo)[1] `

```{r}
########################################### 
## 1
# Select the experiment
i <- 1
p <- pData(geo[[i]][[1]])
head(p)
dis <- "copd status:ch1" 

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("0","1")

new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)


# other infromation
p$PATIENT <- p$`patient:ch1`
p$AGE <- p$`age:ch1`
p$SEX <- p$`Sex:ch1`
p$PACKYEAR <- p$`pack years:ch1`


# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)

# Parse column to have Gene.Symbol
ff <- dplyr::mutate(f,
                   geneInfo = strsplit(gene_assignment, " /// "),
                   gene1 = sapply(geneInfo, `[`, 1),
                   GENE.SYMBOL = sapply(strsplit(gene1, " // "), `[`, 2)
)

rownames(ff) <- rownames(f)
f <- dplyr::select(ff,ID,GB_LIST,GENE.SYMBOL)

length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f

### Adding normalized data 
colnames(norm[[i]]) <- gsub("\\..*","",colnames(norm[[i]]))
exprs(geo[[i]][[1]]) <- exprs(norm[[i]])


########### Summary
df <- obtainInfo(df)

df <- df[-1,]
```


## `r names(geo)[2] `

```{r}
########################################### 
i <- 2
p <- pData(geo[[i]][[1]])
head(p)
dis <- "source_name_ch1"

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("tumor-free tissue of non-smoker",
                "tumor-free tissue of ex-smoker with COPD",
                "tumor-free tissue of smoker with COPD")

new_labels <- c("CONTROL","COPD","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation
p$AGE <- p$`age:ch1`
p$SEX <- p$`gender:ch1`
p$SMOKING_STATUS <- mgsub(p[,dis],old_labels,c("NON-SMOKER","EX-SMOKER","SMOKER"))

# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$GENE_SYMBOL
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f

### Adding normalized data 
colnames(norm[[i]]) <- str_extract(colnames(norm[[i]]),regex("GSM[0-9]*"))
rownames(norm[[i]]) <- norm[[i]]$genes$ProbeName

uni <- unique(norm[[i]]$genes$ProbeName)
uni <- uni[uni %in% rownames(geo[[i]][[1]])]


exprs(geo[[i]][[1]]) <- norm[[i]][uni,]$E


########### Summary
df <- obtainInfo(df)

```


## `r names(geo)[3] `

```{r}
########################################### 
## 3
# Select the experiment
i <- 3
p <- pData(geo[[i]][[1]])
head(p)
dis <- "characteristics_ch1.1"

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("phenotype: healthy smoker","phenotype: Nonsmoker","phenotype: moderate COPD")

new_labels <- c("CONTROL","CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation
p$SMOKING_STATUS <- mgsub(p[,dis],old_labels,c("SMOKER","NON-SMOKER",NA))


# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$`Gene Symbol`
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f

### Adding normalized data 
colnames(norm[[i]]) <- str_extract(colnames(norm[[i]]),regex("GSM[0-9]*"))

exprs(geo[[i]][[1]]) <- exprs(norm[[i]])

########### Summary
df <- obtainInfo(df)

```


## `r names(geo)[4] `

```{r}
library(recount)
## 4
# Select the experiment
i <- 4

## Download data from Recount2
# download_study('SRP041538',outdir = OUTPUT("SRP041538"))
load(file.path(OUTPUT("SRP041538"), 'rse_gene.Rdata'))

## Scale counts by taking into account the total coverage per sample
rse <- scale_counts(rse_gene)
rse

# saving the geoquery object 
geo[[i]][[2]] <- geo[[i]][[1]]

## adding recount2 data to the list 
geo[[i]][[1]] <- rse

### 
p <- colData(geo[[i]][[1]])
head(p)
dis <- "characteristics"

######### DISEASE information
table(unlist(p[,dis]))
ss <- sapply(p[,dis],unlist)
ss <- ss[1,]

# identify the way authors described COPD or CONTROLS
old_labels <- c("disease state: Normal","disease state: COPD")
new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(ss,old_labels,new_labels)
table(p$DISEASE)

# other infromation

# add it to the expression object
colData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- rowData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- as.character(as.data.frame(f)$symbol)
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
rowData(geo[[i]][[1]]) <- f

########### Summary
p <- colData(geo[[i]][[1]])
p <- as.data.frame(p)
pp <- pData(geo[[i]][[2]])

p <- merge(p,pp , by= "geo_accession")

df <- obtainInfo(df)

```

## `r names(geo)[5] `

```{r}
########################################### 
## 5
# Select the experiment
i <- 5
p <- pData(geo[[i]][[1]])
head(p)
dis <- "disease state:ch1" 

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("Control",
                "Chronic Obstructive Lung Disease")
new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation
p$AGE <- p$`age:ch1`
p$SMOKING_STATUS <- p$`smoker?:ch1`

# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$GENE_SYMBOL
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f

### Adding normalized data 
colnames(norm[[i]]) <- str_extract(colnames(norm[[i]]),regex("GSM[0-9]*"))
rownames(norm[[i]]) <- norm[[i]]$genes$ProbeName

uni <- unique(norm[[i]]$genes$ProbeName)
uni <- uni[uni %in% rownames(geo[[i]][[1]])]

exprs(geo[[i]][[1]]) <- norm[[i]][uni,]$E


########### Summary
df <- obtainInfo(df)

```

## `r names(geo)[6] `

```{r}
########################################### 
## 6
# Select the experiment
i <- 6
p <- pData(geo[[i]][[1]])
head(p)
dis <- "diagnosis:ch1" 

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("Normal",
                "COPD")
new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation
p$AGE <- p$`age:ch1`
p$SEX <- p$`gender (1=male, 2=female):ch1`
p$PACKYEAR <- p$`pkyrs:ch1`
p$SMOKING_STATUS <- rep("SMOKER",nrow(p))

# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$GENE_SYMBOL
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f


########### Summary
df <- obtainInfo(df)

```

## `r names(geo)[7] `

```{r}
########################################### 
## 7
# Select the experiment
i <- 7
p <- pData(geo[[i]][[1]])
head(p)

table(gsub('[0-9]+', '', p$title))
p$DISEASE <- "CONTROL"
p$DISEASE[grep("COPD",p$title)] <- "COPD"
p$DISEASE[grep("Unclassified",p$title)] <- "Unclassified"
table(p$DISEASE)


# other infromation
pp <- p$characteristics_ch1

p$AGE <- strapplyc(pp, "Age:.[0-9]*", simplify = TRUE)
p$AGE <- gsub("[^0-9.-]", "", p$AGE)

p$SEX <- strapplyc(pp, "Gender:.[A-Z]*", simplify = TRUE)
p$SEX <-  gsub("[^M|F.-]", "", p$SEX)

# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$`Gene Symbol`
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f

### Adding normalized data 
colnames(norm[[i]]) <- str_extract(colnames(norm[[i]]),regex("GSM[0-9]*"))

exprs(geo[[i]][[1]]) <- exprs(norm[[i]])

########### Summary
df <- obtainInfo(df)


```

## `r names(geo)[8] `

```{r}
########################################### 
## 8
# Select the experiment
i <- 8
p <- pData(geo[[i]][[1]])
head(p)
p$MOD_title <- stringr::str_count(as.character(p$title,"L"))
dis <- "MOD_title" 

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("3",
                "2")
new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation

# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$`Gene Symbol`
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f


### Adding normalized data 
colnames(norm[[i]]) <- str_extract(colnames(norm[[i]]),regex("GSM[0-9]*"))
exprs(geo[[i]][[1]]) <- exprs(norm[[i]])

########### Summary
df <- obtainInfo(df)

```

## `r names(geo)[9] `

```{r}
########################################### 
## 9
# Select the experiment
i <- 9
p <- pData(geo[[i]][[1]])
head(p)
dis <- "description" 

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("Normal lung",
                "'usual' emphysemic lung",
                "Alpha-1 Antitrypsin Deficiency-related emphysemic lung")
new_labels <- c("CONTROL","COPD","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation

# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$`Gene Symbol`
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f


### Adding normalized data 
colnames(norm[[i]]) <- str_extract(colnames(norm[[i]]),regex("GSM[0-9]*"))
exprs(geo[[i]][[1]]) <- exprs(norm[[i]])

########### Summary
df <- obtainInfo(df)

```

## `r names(geo)[10] `


```{r}
########################################### 
## 10
# Select the experiment
i <- 10
p <- pData(geo[[i]][[1]])
head(p)
dis <- "disease state:ch1" 

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("non-emphysema","emphysema")
new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation

# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$`Gene Symbol`
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f


########### Summary
df <- obtainInfo(df)

```



## `r names(geo)[11] `


```{r}
########################################### 
## 11
# Select the experiment
i <- 11
p <- pData(geo[[i]][[1]])
head(p)
dis <- "copd:ch1" 

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("no",
                "yes")
new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation
p$AGE <- p$`age:ch1`
p$SEX <- p$`Sex:ch1`
p$PACKYEAR <- p$`pack years:ch1`
p$SMOKING_STATUS <- p$`smoking:ch1`

# add it to the expression object
pData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$`Gene Symbol`
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]][[1]]) <- f


### Adding normalized data 
colnames(norm[[i]]) <- str_extract(colnames(norm[[i]]),regex("GSM[0-9]*"))

ff <- exprs(norm[[i]])
ff <-  ff[rownames(ff) %in% rownames(f),]
dim(ff)

exprs(geo[[i]][[1]]) <- ff


########### Summary
df <- obtainInfo(df)


```

## `r names(geo)[12] `

```{r}
########################################### 
## 12
# Select the experiment
i <- 12
p <- pData(geo[[i]][[1]])

############ Get gene counts
#getGEOSuppFiles(names(geo)[i])
pp <- read.csv(list.files(OUTPUT("GSE124180"),full.names = T),sep = "\t")
rownames(pp) <- pp$ENSEMBL_GENEID
pp <- pp[,which(colnames(pp)!="X" & colnames(pp)!= "ENSEMBL_GENEID")]
pp <- pp[-1,]
pp <- pp[,which(colnames(pp) %in% p$title)]
colnames(pp[,p$title])

# check order of colnames 
identical(colnames(pp),p$title)

#rename colnames of gene counts df
colnames(pp) <- p$geo_accession

# Get gene annotation
gs <- AnnotationDbi::select(EnsDb.Hsapiens.v79, 
             rownames(pp),
             c("GENEID","GENENAME","SEQNAME","GENESEQSTART","GENESEQEND"), "GENEID")

rownames(gs) <- gs$GENEID

# Select genes with annotation
pp <- pp[which(rownames(pp) %in% gs$GENEID),]

l <- as.matrix(apply(pp, 2, as.numeric))
rownames(l)<- rownames(pp)

#Create the object expressionSet
geo[[i]][[1]] <- SummarizedExperiment(assays=l,
                     rowData=gs, colData=pData(geo[[i]][[1]]))

#### Subset data to  have only lung tissue
p <- colData(geo[[i]][[1]])
table(p$`cell type:ch1`)
geo[[i]][[1]] <- geo[[i]][[1]][,which(p$`cell type:ch1` == "bronchial epithelium")]

p <- colData(geo[[i]][[1]])
head(p)
dis <- "copd:ch1" 

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("cont",
                "case")
new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation
p$AGE <- p$`age:ch1`
p$SEX <- p$`Sex:ch1`
p$PACKYEAR <- p$`pack years:ch1`
p$SMOKING_STATUS <- p$`smoke:ch1`

# add it to the expression object
colData(geo[[i]][[1]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- rowData(geo[[i]][[1]])
head(f)
f$GENE.SYMBOL <- f$GENENAME
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
rowData(geo[[i]][[1]]) <- f


########### Summary

########### Summary
df <- obtainInfo(df)

```


## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>RDS:</strong> ExpressionSet for each experiment with information of samples  
  <strong>Table:</strong> Table of experiment with info summarization
</div>

```{r}

##############################################################
saveRDS(geo,OUTPUT(c(TODAY,"_Step3_LungTissue-CURATED.RDS")))

write_csv(df,OUTPUT(c(TODAY,"_Step3_Summary.csv")))

```



## Session Info

```{r sessionInfo}
sessionInfo()
```





