---
title: "6-meta-analysis.RMD"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To do a meta-analysis for COPD experiments
</div>


We previously had run the following scripts:

- 0: Data selection `/0-Data-selection.RMD`

- 1: Dowload RAW data `/1-download_raw-data.RMD`

- 2: Normalizing data `/2-normalizing-data.RMD`

- 3: Curating data `/3-curatig-data.RMD`

- 4: Differential gene expression analysis  `/4-DE.RMD`

- 5: Merging data `/5-merge-data.RMD`


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Table:</strong> Table with all DE results. Output from step 5   
</div>

## Analysis

### Setup

For running the script, type:

```{bash, eval = F}
ssh -X aaltamirano@dna.lavis.unam.mx
qrsh
cd /mnt/Genoma/amedina/DataDNA/R-projects/Meta-analysis_COPD/output_data/.out
 module load r/4.0.1 

## OR
ssh ana@10.200.0.42
cd /home/ana/DataDNA/R-projects/Meta-analysis_COPD/output_data/.out

nohup R -e "rmarkdown::render(here::here('vignettes/6-meta-analysis.RMD'))" > 6-meta-analysis.RMD.out &
```

The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.


```{r setup}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("Tissue/R/setup.R"))
```

And this analysis is run in: `r getwd() `


### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(ComplexHeatmap)
library(metafor)
library(meta)
library(pheatmap)
library(RColorBrewer)
```


### Read data


```{r experiments}
gse_table <- read.csv(OUTPUT("2020-10-07_Step3_Summary.csv"), row.names = 1)
gse_table <- gse_table[-which(rownames(gse_table)=="GSE119040"),]
kable(gse_table, caption = "GSE information") %>%
  kableExtra::scroll_box(width = "100%", height = "100px")
```


```{r , class.source="bg-info",class.output="bg-info" }
full_tables <- read_csv(OUTPUT("2020-10-08_Step5_Full_Tables.csv"))
colnames(full_tables)

```

### MetaVolcano

Using [Metavolcano](https://www.bioconductor.org/packages/release/bioc/vignettes/MetaVolcanoR/inst/doc/MetaVolcano.html) funcions to create a volcano plot with DE genes 

We first need to create a list with tables. Each table contains information about gene `Symbol`, `Log2FC`, `pvalue`, `CI.L`, `CI.R` per experiment. 

```{r}
# colnames 
df <- tibble(x=colnames(full_tables)) %>% 
  separate(x, c("type", "GSE"), "_") %>% 
  add_column(pos = 1:nrow(.))

df$GSE[1] <- "Gene"

# list
split_df <- function(i){
  df <- data.frame(GENE.SYMBOL = full_tables$GENE.SYMBOL,
                   full_tables[,which(df$GSE == i)]
                   )
  colnames(df) <- c( "Symbol", "Log2FC",  "pvalue", "CI.L", "CI.R")
  #colnames(df) <- gsub("_GSE.*", "", colnames(df) )
  as.data.frame(df)
  #as.data.frame(df)[-c(8359,18704),]
  }

diffexplist <- lapply(unique(df$GSE)[-1], split_df)
names(diffexplist) <- unique(df$GSE)[-1]

```

Now, we implement the `MetaVolcano` funtions. 

```{r}
library(MetaVolcanoR)

meta_degs_rem <- rem_mv(diffexp=diffexplist,
            pcriteria="pvalue",
            foldchangecol='Log2FC', 
            genenamecol='Symbol',
            geneidcol=NULL,
            collaps=FALSE,
            llcol='CI.L',
            rlcol='CI.R',
            vcol=NULL, 
            cvar=TRUE,
            metathr=0.01,
            jobname="MetaVolcano",
            outputfolder=".", 
            draw='HTML',
            ncores=1)

head(meta_degs_rem@metaresult, 3)
meta_degs_rem@MetaVolcano + theme_void()

meta_df <- meta_degs_rem@metaresult %>% 
  filter(randomP < 0.05 )

meta_degs_vote <- votecount_mv(diffexp=diffexplist,
                   pcriteria='pvalue',
                   foldchangecol='Log2FC',
                   genenamecol='Symbol',
                   geneidcol=NULL,
                   pvalue=0.05,
                   foldchange=0, 
                   metathr=0.01,
                   collaps=FALSE,
                   jobname="MetaVolcano", 
                   outputfolder=".",
                   draw='HTML')
head(meta_degs_vote@metaresult, 3)
meta_degs_vote@degfreq
meta_degs_vote@MetaVolcano

```




### p-Values

We first check pvalues histograms. [Ref](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6164648/)

```{r , class.source="bg-info",class.output="bg-info", fig.width= 8 }
FC <- str_subset(colnames(full_tables),"logFC")

# Genes with a adjusted p value
adjp <- full_tables %>%
  filter(GENE.SYMBOL != "NA") %>%
  gather("GSE"  ,"adjP",-GENE.SYMBOL) %>%
  filter(GSE %in% str_subset(colnames(full_tables),"adj.P.Val|padj")) %>%
  mutate(GSE=gsub(".*_","",GSE))
  
adjp


#pdf("Pvalues.pdf",width = 8,height = 6)
adjp %>%    
  group_by(GSE) %>%
  ggplot( aes(x=adjP, fill=GSE)) +
  geom_histogram( color="#e9ecef", alpha=0.6) + #position = 'identity'
  geom_vline(xintercept=0.05, linetype="dashed", color = "red") +
  theme_bw(16) +
  xlab("adjusted p-value") +
  ylab("Frequency") +
  facet_wrap(~GSE, scales = "free_y") 
#dev.off()


```


Then, we look for how many genes are up regulated with a p-value less than `r  adj_pvalue`  in the 

```{r fig.width=10 }

# log fold change
logFC <- full_tables %>%
  filter(GENE.SYMBOL != "NA") %>%
  gather("GSE"  ,"logFC",-GENE.SYMBOL) %>%
  filter(GSE %in% str_subset(colnames(full_tables),"logFC")) %>% 
  mutate(GSE=gsub(".*_","",GSE))

logFC

logFC %>%    
  group_by(GSE) %>%
  ggplot( aes(x=logFC, fill=GSE)) +
  geom_histogram( color="#e9ecef", alpha=0.6) + #position = 'identity'
  geom_vline(xintercept=0, linetype="dashed", color = "red") +
  theme_bw(16) +
  xlab("logFC") +
  ylab("Frequency") +
  facet_wrap(~GSE, scales = "free_x") 

# recovering logFC values from genes with p value less than 0.01 
adj_pvalue <- 0.05


ap <- left_join(adjp,logFC,by = c("GENE.SYMBOL" = "GENE.SYMBOL", "GSE" = "GSE")) %>%
  filter(adjP < adj_pvalue) %>% 
  select(-adjP) %>% 
  spread(GSE,logFC) %>% 
  replace(is.na(.), 0)

ap

# selecting up regulated genes
up <- ap %>%
  mutate_at(vars(-GENE.SYMBOL),list(~if_else( .>0, 1, 0)))
  
down <- ap %>%
  mutate_at(vars(-GENE.SYMBOL),list(~if_else( .<0, 1, 0)))

### UP 
m <- make_comb_mat(up,mode = "distinct")
#m <- m[comb_degree(m) >= 1]
m <- m[comb_degree(m) >= 5 | comb_degree(m) == 1]

#pdf("upsetUP.pdf",width = 14,height = 6)
ht <- draw(UpSet(m,
      pt_size = unit(5, "mm"), 
      lwd = 3,
      comb_col = c("red", "#5c5c5c","#5c5c5c","#5c5c5c", "black","blue")[comb_degree(m)])
      #order.by = "freq"
    )

od = column_order(ht)
cs = comb_size(m)
decorate_annotation("intersection_size", {
    grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(2, "pt"), 
        default.units = "native", just = "bottom", gp = gpar(fontsize = 12))
})

#dev.off()
comb_size(m)
up[extract_comb(m,"11001101"),]
up[extract_comb(m,"10001111"),]


### DOWN
m <- make_comb_mat(down,mode = "distinct")
#m <- m[comb_degree(m) >= 1]
m <- m[comb_degree(m) >= 5 | comb_degree(m) == 1] 

#pdf("upsetDOWN.pdf",width = 14,height = 6)
ht <- draw(UpSet(m,
      pt_size = unit(5, "mm"), 
      lwd = 3,
      comb_col = c("blue", "#5c5c5c","#5c5c5c","#5c5c5c", "black","red")[comb_degree(m)])
      #order.by = "freq"
    )
#dev.off()

od = column_order(ht)
cs = comb_size(m)
decorate_annotation("intersection_size", {
    grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(2, "pt"), 
        default.units = "native", just = "bottom", gp = gpar(fontsize = 12))
})
comb_size(m)
down[extract_comb(m,"10001111"),]
down[extract_comb(m,"01001111"),]


```


## Meta-analysis

We first calculate the Standard error using confidence intervals

```{r}

CI.R <- str_subset(colnames(full_tables),"CI.R")
CI.L <- str_subset(colnames(full_tables),"CI.L")

for(i in 1:length(CI.R)) {
  full_tables[str_c("SE_",gsub(".*_","",CI.R))] <- (full_tables[,CI.R]-full_tables[,CI.L]) /(2*1.96)
}

#write_csv(full_tables,OUTPUT("2020-10-08_Step5_Full_Tables_SE.csv"))
```

Meta-analysis for a single gene


```{r}

FC <- str_subset(colnames(full_tables),"logFC")
SE <- str_subset(colnames(full_tables),"SE_")
GSE <- gsub(".*_","",FC)

gs <- "CXCL12"

## Funtion to perform a meta-analysis per gene
meta.plot <- function(gs){
gene <-  full_tables[full_tables$GENE.SYMBOL == gs,]
#gene <-  na.omit(full_tables[full_tables$GENE.SYMBOL == gs,])
gene <- gene[!is.na(gene$GENE.SYMBOL),]

dat <- data.frame(
  yi = as.numeric(gene[,FC]),
  sei = as.numeric(gene[,SE]),
  n1i = as.numeric(gse_table$CONTROL),
  n2i = as.numeric(gse_table$COPD)
)



dat <- escalc(measure="ZCOR", yi=yi,sei=sei,n1i=n1i,n2i=n2i, data = dat) 

# meta analysis in the simplest way
#res <- metafor::rma(yi, sei=sei,n1i=n1i,n2i=n2i, data=dat)
#funnel(res)
#forest(res,slab = GSE, main=unique(gene$GENE.SYMBOL))

dat$study <- rownames(gse_table)
dat$country <- gse_table$COUNTRY


m1 <- metagen(TE = yi, 
                  seTE = sei, 
                  studlab = study, 
                  n.e = n2i, 
                  n.c = n1i, 
                  data = dat,
                  method.tau = "REML", 
                  prediction = TRUE,
                  sm= "MD",
                  comb.fixed = FALSE)

forest(m1,
       col.diamond = "blue",
       #leftcols = c("study"),
       #rightcols = c("ci", "w.random"),
       leftlabs =  c("Study","logFC","se",NA,NA),
       colgap.left = unit(5,"mm"),
       lab.e = "COPD",
       colgap.studlab = unit(-10,"mm"),
       ff.study.label = "bold",
       just = "center",
       xlab = gs,
       prediction = TRUE)


}

meta.plot(gs)


```

Meta-analysis for all genes. We are saving the results in a new dataframe.

```{r}

#gs <- unique(na.omit(full_tables)$GENE.SYMBOL)
gs <- full_tables$GENE.SYMBOL

meta <- data.frame(genes = 0,
                   TE.random = 0,
                   pval.random = 0,
                   tau2 = 0,
                   I2 = 0,
                   pval.Q = 0,
                   num_exp=0)

# meta <- data.frame(genes = 0,
#                    QEp=0,
#                    estimate=0,
#                    pval=0,
#                    se=0,
#                    H2=0,
#                    tau2=0,
#                    I2=0,
#                    num_exp=0)

gm <- full_tables$GENE.SYMBOL
full_tables[full_tables == 0] <- NA

ft <- full_tables %>% 
  select(GENE.SYMBOL,FC) %>% 
  is.na() %>% 
  rowSums()


length(gm)

for (i in gm) { 
#for( i in gm[18705:length(gm)]){ #problemas con 8359, 18704
  gene <- full_tables[which(full_tables$GENE.SYMBOL == i),]
  dat <- data.frame(
    yi = as.numeric(gene[,FC]),
    sei = as.numeric(gene[,SE]),
    n1i = as.numeric(gse_table$CONTROL),
    n2i = as.numeric(gse_table$COPD)
  )
  dat$study <- rownames(gse_table)
  dat$country <- gse_table$COUNTRY
  

  # meta analysis with random models "REML"
  m1 <- metagen(TE = yi, 
                  seTE = sei, 
                  studlab = study, 
                  n.e = n2i, 
                  n.c = n1i, 
                  data = dat,
                  method.tau = "REML", 
                  prediction = TRUE,
                  sm= "MD",
                  control=list(maxiter=1000), # increasing number of iteration
                  comb.fixed = FALSE)
  # res <- suppressWarnings(metafor::rma(yi, sei=sei,control=list(stepadj=.5,maxiter=1000)))
  num_exp <- sum(!is.na(dat$yi))
  meta <- rbind(meta,
                c(i, m1$TE.random,m1$pval.random,m1$tau2,m1$I2,m1$pval.Q,num_exp))
  #meta <- rbind(meta,c(i,res$QEp,res$beta,res$pval,res$se,res$H2,res$tau2,res$I2,num_exp))
}


# Numeric values
meta <- meta[-1,]
rownames(meta) <- meta$genes

meta <- meta[,-1] %>% mutate_all(as.numeric)

# FDR correction
meta$qval.random <- p.adjust(meta$pval.random,method="fdr")
meta$qval.Q <- p.adjust(meta$pval.Q, method = "fdr")

#write.csv(meta, OUTPUT(c(TODAY,"_Step6_meta-analysis.csv")))
meta <- read_csv("Tissue/output_data/2021-03-01_Step6_meta-analysis.csv")

### Distribution of statistics 

bqval <- ggplot(meta) +
 aes(x = qval.random) +
 geom_histogram(bins = 55L, fill = "#5abdc4") +
 labs(x = "q-value from REML") +
 theme_classic()

bq <- ggplot(meta) +
 aes(x = qval.Q) +
 geom_histogram(bins = 55L, fill = "#5abdc4") +
 labs(x = "q-value for Q test") +
 theme_classic()

bte <- ggplot(meta) +
 aes(x = TE.random) +
 geom_histogram(bins = 55L, fill = "#5abdc4") +
 labs(x = "Estimate values") +
 theme_classic()

btau <- ggplot(meta) +
 aes(x = tau2) +
 geom_histogram(bins = 55L, fill = "#5abdc4") +
 labs(x = "Tau2") +
 theme_classic()

bi2 <- ggplot(meta) +
 aes(x = I2) +
 geom_histogram(bins = 55L, fill = "#5abdc4") +
 labs(x = "I2") +
 theme_classic()


ggpubr::ggarrange(bqval,bq,bte,btau,bi2,
                  nrow =2,
                  ncol =3)

### Selecting genes
qval_cutoff <- 0.05

genesqval <- meta %>% 
  #filter(meta$qval.random <= qval_cutoff ) %>% 
  filter(meta$qval.random <= qval_cutoff & I2 < 0.40 & num_exp >= 9) %>% 
  arrange(TE.random)

genesqval

kable(c(head(genesqval),tail(genesqval)), caption = "Meta-analysis")

### Forest plot
meta.plot(genesqval$X1[1])
meta.plot(genesqval$X1[nrow(genesqval)])

### logFC values from selected genes
topgenes <- full_tables %>% 
  filter(full_tables$GENE.SYMBOL %in% genesqval$X1) %>% 
  select(GENE.SYMBOL,contains("logFC"))

## as dataframe
topgenes <- column_to_rownames(topgenes,"GENE.SYMBOL")

# scaling data
topgenes <- scale(topgenes)


#"BrBG"
color <- rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(15))
breaks <- c(seq(min(topgenes,na.rm = T),-0.1,length.out = 7),0,seq(0.1,max(topgenes,na.rm = T),length.out = 7))




```

```{r fig.height=15}

dim(topgenes)
pheatmap(topgenes,
         main = str_c("Genes with q-value less than ",qval_cutoff, " (ngenes = ",dim(topgenes)[1], ")"),
         color = color,
         breaks = breaks,
         fontsize_row = 7,
         fontsize_col = 10,
         cluster_rows = T,
         clustering_distance_rows = "euclidean",
         clustering_method = "ward.D",
         cluster_cols = T,
         cutree_rows = 2,
         show_colnames = T,
         show_rownames = F,
         border_color = NA)

```


```{r}
mo <-  genesqval %>% 
  arrange(TE.random)

mo <- rbind(head(mo,15), tail(mo,15))

### logFC values from selected genes
topgenes <- full_tables %>% 
  filter(full_tables$GENE.SYMBOL %in% mo$X1) %>% 
  select(GENE.SYMBOL,contains("logFC")) %>% 
  column_to_rownames(.,"GENE.SYMBOL") 

topgenes <- scale(topgenes)

color <- rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(15))
breaks <- c(seq(min(topgenes,na.rm = T),-0.1,length.out = 7),0,seq(0.1,max(topgenes,na.rm = T),length.out = 7))

pheatmap(topgenes,
         main = str_c("The highest logFC and qval < " ,qval_cutoff , " (ngenes = ",dim(topgenes)[1], ")"),
         color = color,
         breaks = breaks,
         fontsize_row = 9,
         fontsize_col = 10,
         cluster_rows = T,
         clustering_distance_rows = "euclidean",
         clustering_method = "ward.D",
         cluster_cols = T,
         cutree_rows = 2,
         show_colnames = T,
         border_color = NA)

```


### TFs

We also checked for expression in TFs 

````{r}
tfs <- read.csv(here::here("Tissue/data/hs_hgnc_curated_tfs.txt"),col.names = "TF")

tf_copd <- filter(meta,meta$X1 %in% tfs$TF)
tf_copd <- filter(genesqval,genesqval$X1 %in% tfs$TF)
#write.csv(tf_copd, OUTPUT(c(TODAY,"_Step6_TF-",qval_cutoff,"-meta-analysis.csv")))

topgenes <- full_tables %>% 
  filter(full_tables$GENE.SYMBOL %in% tf_copd$X1) %>% 
  select(GENE.SYMBOL,contains("logFC")) %>% 
  column_to_rownames("GENE.SYMBOL")

topgenes <- scale(topgenes)
#topgenes <- na.omit(topgenes)


color <- rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(15))
breaks <- c(seq(min(topgenes,na.rm = T),-0.1,length.out = 7),0,seq(0.1,max(topgenes,na.rm = T),length.out = 7))

pheatmap(na.omit(topgenes),
         main = str_c("TFs" , " (ngenes = ",dim(topgenes)[1], ")"),,
         color = color,
         breaks = breaks,
         fontsize_row = 10,
         fontsize_col = 10,
         cluster_rows = T,
         clustering_distance_rows = "euclidean",
         clustering_method = "ward.D",
         cluster_cols = T,
         cutree_rows = 2,
         show_colnames = T,
         border_color = NA)

```



## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Plots:</strong> Upset plots, forest plot, heatmap, etc.     
  <strong>Table:</strong> Results from the meta-analysis `r OUTPUT(c(TODAY,"_Step6_meta-analysis.csv"))` 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





