---
title: "2 Blood: Normalizing data"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    self_contained: no
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To normalize raw data
</div>

### Background 

Some relevant information with links [GEO](https://www.ncbi.nlm.nih.gov/geo/)


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Data:</strong> Raw data (e.i. .CEL, .TXT)  
  <strong>2-Table_GSE-info.txt:</strong> Table with GSE IDs and platform description 
</div>



## Setup

For running the script, type:

```{bash, eval = F}
ssh -X aaltamirano@dna.lavis.unam.mx
qrsh
cd /home/ana/DataDNA/R-projects/Meta-analysis_COPD/Blood/output_data/.out
 module load r/4.0.1 

nohup R -e "rmarkdown::render(here::here('Blood/vignettes/2-Blood-normalizing-data.RMD'),output_options = c('self_contained = TRUE'))" > 2-Blood-normalizing-data.out &

```

The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.

```{r setup,}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("Blood/R/setup.R"))
```

And this analysis is run in: `r getwd() `


### Libraries

```{r libraries, message=FALSE, warning=FALSE,eval = T}
library(limma)
library(oligo)
library(tidyverse)
```


## Functions


### **rawCEL_normCEL**: Pre-process raw data for Affymetrix 

We first read and pre-process raw *.CEL* files. Each experiment has it own 
folder with raw data per sample. Then we normalized using RMA algorithm and 
finally, we save it. 

Be sure files are untar and unzip.

In this function you need:

<div class="alert alert-warning">
  <strong>Input:</strong> GSE ID   
  <hr>
  <strong>Output:</strong>  
  - Samples normalized  
  - Boxplots  
  - Histograms 
</div>


```{r rawCEL_normCEL,eval = T}
rawCEL_normCEL <- function(gse){
  
  # select and untar files
  gsepath <- file.path(here::here("GEOquery"),gse)
  print(gse)
  
  if(length(list.files(gsepath)) == 0){
    message("No Supplementary files found")
    } else if (length(list.files(gsepath)) < 3 & length(list.files(gsepath)) != 0 ){
    ifelse(any(list.files(gsepath,pattern="tar") !=0),
                untar(list.files(gsepath,full.names=T,pattern="tar"), exdir=gsepath),
                list.files(gsepath))
    } else{
    print(length(list.files(gsepath)))
    }
  
  # for CEL files "AFFYMETRIX"
  if(any(grepl("cel",list.files(gsepath),ignore.case = T))){
    message("This is an Affymetrix array")
    celfiles <- list.celfiles(gsepath,
                              full.names=TRUE,
                              listGzipped=TRUE)
    
    # read CEL files in R
    rawData <- read.celfiles(celfiles)
    #### Figures of raw data
    #pdf(str_c("raw_",gse,"_boxplot",TODAY,".pdf"))
    ## plots of raw data
    boxplot(rawData,target="core")
    hist(rawData,target="core")
    #dev.off()
    ## RMA normalization
    normData <- rma(rawData)
    #### Figures of Normalized data
    #pdf(str_c("norm_",gse,"_boxplot",TODAY,".pdf"))
    ## `plots of norm data
    boxplot(normData)
    hist(normData)
    # dev.off()
    #write.csv(exprs(normData),str_c("/normalized",gse,"_normData",TODAY,".tx  t"),quote=F)
    return(normData)
  }
  
  #for TXT files "AGILENT"
  else if(sum(grepl("txt",list.files(gsepath),ignore.case = T)) >2){
    message("This is an Agilent array")
    files <- list.files(gsepath,full.names=T, pattern = "txt")
    files <- grep("annot",files,value=T,invert=T)
   
     # read agilent files
    RG <- read.maimages(files,source = "agilent.median",green.only=T)
    
    #plots: raw data
     boxplot(RG$E)
     hist(RG$E)
    # normalizing data
    RG <- limma::backgroundCorrect(RG, method="normexp", offset=1)
    RG$E <- normalizeBetweenArrays(RG$E, method="quantile")
    RG$E <- log2(RG$E)
    #plots: norm
    boxplot(RG$E)
    hist(RG$E,100)
    
    return(RG)
  }
  
  #for TXT files "GPR"
  else if(sum(grepl("gpr",list.files(gsepath),ignore.case = T)) >2){
    message("This is an Agilent array with GPR files")
    files <- list.files(gsepath,full.names=T, pattern = "gpr")
    files <- grep("annot",files,value=T,invert=T)
   
     # read agilent files
    RG <- read.maimages(files,source = "genepix",green.only=T)
    #plots: raw data
     boxplot(RG$E)
     hist(RG$E)
    # normalizing data
    RG <- limma::backgroundCorrect(RG, method="normexp", offset=1)
    RG$E <- normalizeBetweenArrays(RG$E, method="quantile")
    RG$E <- log2(RG$E)
    #plots: norm
    boxplot(RG$E)
    hist(RG$E,100)
    
    return(RG)
  }
  # anyother case
  else{
    message("This is not Affymetrix or Agilent")
  }
}

#sapply(tissue,rawCEL_normCEL)

```


## Analysis

We had selected this experiments and downloaded raw data using srcripts:

- 0: Data selection `/0-Data-selection.RMD`

- 1: Dowload RAW data `/1-download_raw-data.RMD`


```{r}
blood <- scan(DATA("blood_GSE_IDs.txt"),
               character(), 
               sep = ",")
```

### STEP 1: Normalizing raw data

Normalize raw data

```{r,eval = T}
geo_norm <- sapply(blood,rawCEL_normCEL)
saveRDS(geo_norm,OUTPUT(c(TODAY,"_normData.xz")),compress = "xz")
```


## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>`2020-09-12_normData.RDS`:</strong> Normalized data 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





