---
title: "3-Blood-curating-data.RMD"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To manually curate sample information and to join meta data with normalized data
</div>

We previously had run the following scripts:


- 1: Dowload RAW data `/1-Blood-download_raw-data.RMD`

- 2: Normalizing data `/2-Blood-normalizing-data.RMD`

## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Data 1:</strong> Data information from geo Blood_2020-11-24_GEOquery-download.RDS     
  <strong>Data 2:</strong> Matrix of normalized experiments Blood_2020-11-24_normData.xz  
</div>


## Setup

```{bash, eval = F}
ssh -X aaltamirano@dna.lavis.unam.mx
qrsh
cd /home/ana/DataDNA/R-projects/Meta-analysis_COPD/Blood/output_data/.out
 module load r/4.0.1 

nohup R -e "rmarkdown::render(here::here('Blood/vignettes/3-Blood-curatig-data.RMD'))" > 3-Blood-curatig-data.RMD.out &

```

The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.

```{r setup,}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("Blood/R/setup.R"))
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(GEOquery)
library(mgsub)
library(annotate)
library(hgu133plus2.db)
library(tidyverse)
library(gsubfn)
library(affycoretools)
```


## Curation

### STEP 1: Import data

We read GEOquery data for all the experiments, it has the 

```{r}
geo <- readRDS(OUTPUT("2020-11-24_GEOquery-download.RDS"))
norm <- readRDS(OUTPUT("2020-11-24_normData.xz"))

names(geo)
names(norm)
```


### STEP 2: Curate information


**NEW COLUMN and CONTROLED VOCABULARY!!**

We will add a new column and use controlled vocabulary, so we will change the 
description that each authors used to describe COPD or Control samples.

In addition, we will create a new table summarizing data from all experiments, 
it will contain number of COPD samples, controls, experiment publication date, etc. 

```{r}
######## SUMMARY TABLE
## Summary information about the experiment
df <- tibble("GSE"= NA,
             "CONTROL" = NA,
             "COPD" = NA,
             "COUNTRY" = NA,
             "SUBMISSION_DATE" = NA,
             "PLATFORM" = NA,
             "NUM_GENES" = NA,
             "AGE_MEAN" = NA,
             "AGE_MEDIAN" = NA,
             "AGE_RANGE" = NA,
             "SEX_F" = NA,
             "SEX_M" = NA,
             "SMOKING_STATUS" = NA,
             "PACKYEAR" = NA
                 )

```


## `r names(norm)[1] `

This experiment has healthy volunteers (HV) with pre and post inhaled lipopolysaccharide (LPS)
treatment, in this analysis we will only use pre treatment as a control group.

```{r}
########################################### 
## 1
# Select the experiment
i <- 1
p <- pData(geo[[i]])
head(p)
dis <- "disease:ch1" 
tre <- "challenge_time_point:ch1"

######### DISEASE information
table(p[,dis])
table(p[,tre])

# identify the way authors described COPD or CONTROLS
old_labels <- c("HVT","COPD")
new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
# identifying controls post treatment
p$DISEASE[grep("pre",p[,tre])] <- "post-LPS-CONTROL"

table(p$DISEASE)


# other infromation
NA


# add it to the expression object
pData(geo[[i]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]])
f[1,]

# Parse column to have Gene.Symbol
ff <- dplyr::mutate(f,
                   geneInfo = strsplit(`Gene Symbol`, " /// "),
                   GENE.SYMBOL = sapply(geneInfo, `[`, 1),
)

rownames(ff) <- rownames(f)
f <- dplyr::select(ff,ID,GB_ACC,ENTREZ_GENE_ID,GENE.SYMBOL)

length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]]) <- f

### Adding normalized data 
colnames(norm[[i]]) <- gsub("_..*","",colnames(norm[[i]]))
exprs(geo[[i]]) <- exprs(norm[[i]])


########### Summary
df <- df %>% add_row(
        "GSE" = names(norm)[i],
        "CONTROL" = table(p$DISEASE)[1],
        "COPD" = table(p$DISEASE)[2],
        "COUNTRY" = unique(p$contact_country),
        "SUBMISSION_DATE" = paste(unique(p$submission_date),collapse = ","),
        "PLATFORM" = unique(p$platform_id),
        "NUM_GENES" = length(unique(f$GENE.SYMBOL))
)

df <- df[-1,]
```


## `r names(norm)[2] `

This experiment has anti-CD3/anti-CD28 treatment but in this analysis those 
samples were not use. 

```{r}
########################################### 
i <- 2
p <- pData(geo[[i]])
head(p)
dis <- "diagnosis:ch1"

# Identify stimulated samples
p$Tre <- gsub('_[0-9]+', '', p$title)

######### DISEASE information
table(p[,dis])
table(p$Tre)

# identify the way authors described COPD or CONTROLS
old_labels <- c("Never Smoker",
                "Smoker",
                "COPD",
                "Screen-fail")

new_labels <- c("CONTROL","CONTROL","COPD","Screen-fail")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)

# changing label for stimulated samples
p$DISEASE[grep("stimulated whole blood",p$Tre)] <- "stimulated-SAMPLES"
table(p$DISEASE)

# other infromation
p$SMOKING_STATUS <- mgsub(p[,dis],old_labels,c("NON-SMOKER","SMOKER","NA","NA"))
table(p$SMOKING_STATUS)

# add it to the expression object
pData(geo[[i]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]])
f[1,]

# Parse column to have Gene.Symbol
ff <- dplyr::mutate(f,
                   geneInfo = strsplit(`Gene Symbol`, " /// "),
                   GENE.SYMBOL = sapply(geneInfo, `[`, 1),
)

rownames(ff) <- rownames(f)
f <- dplyr::select(ff,ID,GB_ACC,ENTREZ_GENE_ID,GENE.SYMBOL)

length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]]) <- f

### Adding normalized data 
colnames(norm[[i]]) <- gsub("_..*","",colnames(norm[[i]]))
exprs(geo[[i]]) <- exprs(norm[[i]])

########### Summary
df <- df %>% add_row(
        "GSE" = names(norm)[i],
        "CONTROL" = table(p$DISEASE)[1],
        "COPD" = table(p$DISEASE)[2],
        "COUNTRY" = unique(p$contact_country),
        "SUBMISSION_DATE" = paste(unique(p$submission_date),collapse = ","),
        "PLATFORM" = unique(p$platform_id),
        "NUM_GENES" = length(unique(f$GENE.SYMBOL)),
        "SMOKING_STATUS" = toString(c(names(table(p$SMOKING_STATUS)),table(p$SMOKING_STATUS)))
)


```


## `r names(norm)[3] `

```{r}
########################################### 
## 3
# Select the experiment
i <- 3
p <- pData(geo[[i]])
head(p)
dis <- "finalgold:ch1"

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("0","1","2","3","4","-1")
new_labels <- c("CONTROL",rep("COPD",5))

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# other infromation
p$SMOKING_STATUS <- p$`ats_packyears:ch1`
p$AGE <- p$`age_enroll:ch1`
p$SEX <- p$`gender:ch1`

########### Gene.Symbol
norm[[i]] <- annotateEset(norm[[i]], hgu133plus2.db)
fData(norm[[i]])$GENE.SYMBOL <- fData(norm[[i]])$SYMBOL

# add it to the expression object
colnames(norm[[i]]) <- gsub("_..*","",colnames(norm[[i]]))
pData(norm[[i]]) <- p

geo[[i]] <- norm[[i]]



########### Summary
df <- df %>% add_row(
        "GSE" = names(norm)[i],
        "CONTROL" = table(p$DISEASE)[1],
        "COPD" = table(p$DISEASE)[2],
        "COUNTRY" = unique(p$contact_country),
        "SUBMISSION_DATE" = paste(unique(p$submission_date),collapse = ","),
        "PLATFORM" = unique(p$platform_id),
        "NUM_GENES" = length(unique(f$GENE.SYMBOL)),
        "AGE_MEAN" = mean(as.numeric(p$AGE),na.rm = T),
        "AGE_MEDIAN" = median(as.numeric(p$AGE),na.rm = T),
        "AGE_RANGE" = paste(as.character(range(as.numeric(p$AGE),na.rm = T)),collapse= "-"),
        "SEX_F" = p$SEX[1],
        "SEX_M" = p$SEX[2],
        "PACKYEAR" =  median(na.omit(as.numeric(p$PACKYEAR)))
)


```


## `r names(norm)[4] `

```{r}
########################################### 
## 4
# Select the experiment
i <- 4
p <- pData(geo[[i]])
head(p)
dis <- "source_name_ch1" 

######### DISEASE information
table(p[,dis])

# identify the way authors described COPD or CONTROLS
old_labels <- c("Health","COPD")
new_labels <- c("CONTROL","COPD")

# rename 
p$DISEASE <- mgsub(p[,dis],old_labels,new_labels)
table(p$DISEASE)

# add it to the expression object
pData(geo[[i]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]])
head(f)
f$GENE.SYMBOL <- f$GENE_SYMBOL
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]]) <- f

### Adding normalized data 
colnames(norm[[i]]) <- str_extract(colnames(norm[[i]]),regex("GSM[0-9]*"))
exprs(geo[[i]]) <- norm[[i]]$E


########### Summary
df <- df %>% add_row(
        "GSE" = names(norm)[i],
        "COPD" = table(p$DISEASE)[2],
        "COUNTRY" = unique(p$contact_country),
        "SUBMISSION_DATE" = paste(unique(p$submission_date),collapse = ","),
        "PLATFORM" = unique(p$platform_id),
        "NUM_GENES" = length(unique(f$GENE.SYMBOL))
)


```

## `r names(norm)[5] `

```{r}
########################################### 
## 5
# Select the experiment
i <- 5
p <- pData(geo[[i]])
head(p)

######### DISEASE information
# Selecting COPD samples
p$DISEASE <- "NA"
p$DISEASE[which(p$`fev1fvc:ch1` < 70)] <- "COPD"
p$DISEASE[which(p$`fev1fvc:ch1` >= 70)] <- "CONTROL"

table(p$DISEASE)

# other infromation
p$AGE <- p$`age:ch1`
p$SEX <- p$`gender:ch1`
p$PACKYEAR <- p$`packs:ch1`


# add it to the expression object
pData(geo[[i]]) <- p

########### Gene.Symbol
# We also add a new column with Gene.Symbol annotation
f <- fData(geo[[i]])
head(f)
f$GENE.SYMBOL <- f$`Gene Symbol`
length(f$GENE.SYMBOL)
length(unique(f$GENE.SYMBOL))
fData(geo[[i]]) <- f


########### Summary
df <- df %>% add_row(
        "GSE" = names(norm)[i],
        "CONTROL" = table(p$DISEASE)[1],
        "COPD" = table(p$DISEASE)[2],
        "COUNTRY" = unique(p$contact_country),
        "SUBMISSION_DATE" = paste(unique(p$submission_date),collapse = ","),
        "PLATFORM" = unique(p$platform_id),
        "NUM_GENES" = length(unique(f$GENE.SYMBOL)),
        "AGE_MEAN" = mean(as.numeric(p$AGE),na.rm = T),
        "AGE_MEDIAN" = median(as.numeric(p$AGE),na.rm = T),
        "AGE_RANGE" = paste(as.character(range(as.numeric(p$AGE),na.rm = T)),collapse= "-"),
        "SEX_F" = p$SEX[1],
        "SEX_M" = p$SEX[2],
        "PACKYEAR" =  median(na.omit(as.numeric(p$PACKYEAR)))
)

```


## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>RDS:</strong> ExpressionSet for each experiment with information of samples  
  <strong>Table:</strong> Table of experiment with info summarization
</div>

```{r}

##############################################################
saveRDS(geo,OUTPUT(c(TODAY,"_Step3_LungTissue-CURATED.RDS")))

write_csv(df,OUTPUT(c(TODAY,"_Step3_Summary.csv")))

```



## Session Info

```{r sessionInfo}
sessionInfo()
```





