---
title: "5-Blood-merge-data.RMD"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To merge data from individual DE analysis
</div>

We previously had run the following scripts:

- 1: Dowload RAW data `/1-Blood-download_raw-data.RMD`

- 2: Normalizing data `/2-Blood-normalizing-data.RMD`

- 3: Curating data `/3-Blood-curatig-data.RMD`

- 4: Differential gene expression analysis  `/4-Blood-DE.RMD`


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Data:</strong> Tables of DE results in a `csv` files  
</div>

## Analysis

### Setup

```{bash, eval = F}
ssh -X aaltamirano@dna.lavis.unam.mx
qrsh
cd /home/ana/DataDNA/R-projects/Meta-analysis_COPD/Blood/output_data/.out
 module load r/4.0.1 

nohup R -e "rmarkdown::render(here::here('Blood/vignettes/5-Blood-merge-data.RMD'))" > 5-Blood-merge-data.RMD &

```

The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.

```{r setup,}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("Blood/R/setup.R"))
```

And this analysis is run in: `r getwd() `


### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
```


### Import data

We will import DE results from all experiments. 

```{r}
files <- list.files(here::here("Blood/output_data/DE"), full.names = T)
files <- gtools::mixedsort(files)[-10]
data <- files %>% map(read_csv)
```


### Merge data

We will use a function to select columns we already curated 

Using `column_selection` function, we can subset columns that has Gene symbol,
logFC, adjsted p-value, and confidence intervals with unique id. 

```{r column_selection, class.source="bg-info",class.output="bg-info" }

column_selection <- function(df){
  # Compete column name (i.e "Gene.Symbol_GSE1122")
  GeneS <- str_subset(colnames(df),"GENE[.]SYMBOL")
  logFC <- str_subset(colnames(df),"logFC")
  adjP <- str_subset(colnames(df),"adj.P.Val|padj")
  CI.L <- str_subset(colnames(df),"CI.L")
  CI.R <- str_subset(colnames(df),"CI.R")
  
  selectedColumns <- c(GeneS,logFC,adjP,CI.L,CI.R)
  #Return subset of dataframe to only keep the first duplicate
  DF <- df[!duplicated(df[,GeneS]),selectedColumns] %>%
    rename(GENE.SYMBOL=GeneS)
  
  return(DF)
}


```


Join all data into one dataset!

```{r merge}
full_tables <- data %>% 
  map(column_selection)%>% 
  purrr::reduce(full_join, by = "GENE.SYMBOL") 

full_tables  

## Select genes with at least 3 experiments with data information
genes <- full_tables %>% 
  select(GENE.SYMBOL,str_subset(colnames(full_tables),"logFC")) %>%
  filter(rowSums(is.na(.)) < 9 ) %>% # at least 3 experiments with logFC (9/12 NA)
  select(GENE.SYMBOL)

full_tables <- filter(full_tables, GENE.SYMBOL %in% genes$GENE.SYMBOL)
full_tables <- full_tables[which(!is.na(full_tables$GENE.SYMBOL)),]

full_tables

## Save data 
write_csv(full_tables, file= OUTPUT(c(TODAY,"_Step5_Full_Tables.csv")))
```


## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Table:</strong> All DE joined using the same gene ID  `r OUTPUT(c(TODAY,"_Step5_Full_Tables.csv"))`  
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





