---
title: "7-Blood-enrichment-analysis.RMD"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To merge data from individual DE analysis
</div>

We previously had run the following scripts:

- 1: Dowload RAW data `/1-Blood-download_raw-data.RMD`

- 2: Normalizing data `/2-Blood-normalizing-data.RMD`

- 3: Curating data `/3-Blood-curatig-data.RMD`

- 4: Differential gene expression analysis  `/4-Blood-DE.RMD`

- 5: Merging data `/5-Blood-merge-data.RMD`

- 6: Merging data `/6-Blood-meta-analysis.RMD`




## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Data:</strong> Meta-analysis results in `csv` files  
</div>

## Analysis

### Setup

For running the script, type:


```{bash, eval = F}
ssh -X aaltamirano@dna.lavis.unam.mx
qrsh
cd /home/ana/DataDNA/R-projects/Meta-analysis_COPD/Blood/output_data/.out
 module load r/4.0.1 

nohup R -e "rmarkdown::render(here::here('Blood/vignettes/7-Blood-enrichment-pathways.RMD'))" > 7-Blood-enrichment-pathways.RMD &

```


The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.


```{r setup}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("Blood/R/setup.R"))
```

And this analysis is run in: `r getwd() `


### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(hypeR)
library(WGCNA)
library(tidyverse)
library(reactable)
```


### Import data

We will import DE results from all experiments. 

```{r}
meta <- read_csv(OUTPUT("2020-11-30_Step6_meta-analysis.csv"))[,-1]
meta
```



### Ranking and weighted values

```{r , class.source="bg-info",class.output="bg-info" }
qval_cutoff <- 0.5

df <- meta %>%
  filter(qval <= qval_cutoff & meta$num_exp > 2)  %>%
  dplyr::arrange(desc(estimate))

signature <- as.vector(df$estimate)
names(signature) <- df$genes
head(signature)
```

Then we will check KEGG, REACTOME and some databases from EnrichR


```{r, eval=FALSE}
# KEGG
KEGG <- hyperdb_rgsets("KEGG", "92.0")

#HALLMARL
HALLMARK <- msigdb_gsets(species="Homo sapiens", category="H",clean = T)

#ENRICHR
TF_TJ <- enrichr_gsets("TRANSFAC_and_JASPAR_PWMs")

enrichR<- enrichr_gsets("Jensen_DISEASES")

enrichR<- enrichr_gsets("Jensen_TISSUES")

enrichR<- enrichr_gsets("Chromosome_Location")

enrichR<- enrichr_gsets("WikiPathways_2019_Human")

enrichR<- enrichr_gsets("ClinVar_2019")

```



```{r}
KEGG <- hyperdb_rgsets("KEGG", "92.0")

DB <- KEGG
  # Ranked Signature using the estimator from the meta analysis (FC)
  hyp_obj <- hypeR(names(signature),DB, test="hypergeometric", background=nrow(meta), fdr=0.5, plotting=TRUE)
  hyp_obj$plots[[1]]
  hyp_obj
  
  # Weighted Signature using as ranking the estimator and the weighted the qvalue
  hyp_obj <- hypeR(signature, DB, test="kstest", background=nrow(meta),fdr=0.5, plotting=TRUE)
  hyp_obj$plots[[1]]
  hyp_obj
  
  # Show interactive table
  #hyp_show(hyp_obj)
  
  # Plot dots plot
  hyp_dots(hyp_obj) + theme_minimal()
  
  # Plot enrichment map
  #hyp_emap(hyp_obj)
  
  # Plot hiearchy map
  #hyp_hmap(hyp_obj)
  
  # Generate markdown report
  #hyp_to_rmd(hyp_obj,"hypr-KEGG-genesets.rmd")
  #hyp_to_rmd(hyp_obj,"hypr-TRANSFAC_and_JASPAR_PWMs.rmd")
```




```{r}
HALLMARK <- msigdb_gsets(species="Homo sapiens", category="H",clean = T)
DB <- HALLMARK
  # Ranked Signature using the estimator from the meta analysis (FC)
  hyp_obj <- hypeR(names(signature),DB, test="hypergeometric", background=nrow(meta), fdr=0.5, plotting=TRUE)
  hyp_obj$plots[[1]]
  hyp_obj
  
  # Weighted Signature using as ranking the estimator and the weighted the qvalue
  hyp_obj <- hypeR(signature, DB, test="kstest", background=nrow(meta),fdr=0.5, plotting=TRUE)
  hyp_obj$plots[[1]]
  hyp_obj
  
  # Show interactive table
  #hyp_show(hyp_obj)
  
  # Plot dots plot
  hyp_dots(hyp_obj) + theme_minimal()
  
  # Plot enrichment map
  #hyp_emap(hyp_obj)
  
  # Plot hiearchy map
  #hyp_hmap(hyp_obj)
  
  # Generate markdown report
  #hyp_to_rmd(hyp_obj,"hypr-KEGG-genesets.rmd")
  #hyp_to_rmd(hyp_obj,"hypr-TRANSFAC_and_JASPAR_PWMs.rmd")
```


```{r}
TF_TJ <- enrichr_gsets("TRANSFAC_and_JASPAR_PWMs")
DB <- TF_TJ
  # Ranked Signature using the estimator from the meta analysis (FC)
  hyp_obj <- hypeR(names(signature),DB, test="hypergeometric", background=nrow(meta), fdr=0.5, plotting=TRUE)
  hyp_obj$plots[[1]]
  hyp_obj
  
  # Weighted Signature using as ranking the estimator and the weighted the qvalue
  hyp_obj <- hypeR(signature, DB, test="kstest", background=nrow(meta),fdr=0.5, plotting=TRUE)
  hyp_obj$plots[[1]]
  hyp_obj
  
  # Show interactive table
  #hyp_show(hyp_obj)
  
  # Plot dots plot
  hyp_dots(hyp_obj) + theme_minimal()
  
  # Plot enrichment map
  #hyp_emap(hyp_obj)
  
  # Plot hiearchy map
  #hyp_hmap(hyp_obj)
  
  # Generate markdown report
  #hyp_to_rmd(hyp_obj,"hypr-KEGG-genesets.rmd")
  #hyp_to_rmd(hyp_obj,"hypr-TRANSFAC_and_JASPAR_PWMs.rmd")
```





## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Figures in the vignette:</strong> Enrichment analysis with plots
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





