---
title: "0: Data selection"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

<div class="alert alert-info">
  <strong>Aim:</strong> To select datasets that measure gene expression in COPD patients. 
</div>

# Introduction

Selecting the experiments is a key point in a meta-analysis so it is important 
to decide which experiments will be used in the analysis. Raw and normalized 
data from gene expression analysis are stored in public repositories. [GEO](https://www.ncbi.nlm.nih.gov/geo/) from 
NCBI and [ArrayExpress](https://www.ebi.ac.uk/arrayexpress/) 
from EMBL-EBI are the most popular, the data is available for everyone and both 
databases have web interface to search and download data.


## Search of papers related with COPD and transcriptomics in GEO

Using `e-utilities` from [NCBI](https://www.ncbi.nlm.nih.gov/books/NBK179288/?report=classic),
I searched articles related with 'Chronic Obstructive Pulmonary Disease' or 'COPD' and
'microarrays' or 'RNA sequencing'. 


```{bash eval= F}
# Search for articles in pubmed
 esearch -db pubmed -query '(("chronic obstructive pulmonary disease" OR "COPD") AND ("microarrays" OR ("RNA sequencing"))' |
  efetch -format docsum |
  xtract -pattern DocumentSummary -element Id PubDate Authors Title FullJournalName > output_data/pubmed_articles.txt
  
# Get abstracts in pubmed
 esearch -db pubmed -query '(("chronic obstructive pulmonary disease" OR "COPD") AND ("microarrays" OR ("RNA sequencing"))' |
  efetch -format abstract > output_data/pubmed_abstract.txt
```


```{r}
pubmed <- read.csv(here::here("output_data/pubmed_articles.txt"), sep = "\t", header = F)
head(pubmed)
```


### Retrieve GSEs 

We also use `e-utilities` from [NCBI](https://www.ncbi.nlm.nih.gov/books/NBK179288/?report=classic) to retrieve GSE ids that are related with COPD from human samples.


```{bash eval= F}
# Search for GSE in NCBI
esearch -db gds -query '(("chronic obstructive pulmonary disease" OR "COPD" OR "Emphysema" OR "chronic bronchitis") AND ("Expression profiling by high throughput sequencing"[Filter] OR "Expression profiling by array"[Filter]) AND "Homo sapiens"[orgn] AND "gse"[Filter]' |
  efetch -format xml >  output_data/GSE_query.txt

perl scripts/parsegeo.pl output_data/GSE_query.txt
```

Table looks like this

```{r}
###################################     
# Read table "output_data/GSE_query_tab.txt"
###################################

gse <- read.csv(here::here("GEO_data/GSE_query_tab.txt"), 
                sep = "\t", 
                header = T,
                row.names = 1)

head(gse)

###################################     
# Libraries
###################################

library(stringr)
library(tidyr)

###################################     
# find_text() function
###################################
# function to find words in the description and title columns
# and add a string in a new vector
find_text <- function(text){
gse$x <- NA
gse$x[str_which(tolower(str_c(gse$Description,gse$Extended.Description)), tolower(text))] <- toupper(text)
return(gse$x)
}



###################################
# "SampleType" as new column
###################################
# Blood related 
gse$Blood <- find_text("Blood")
# Tissue related
gse$Tissue <- find_text("Tissue")
# Airway Epithelium 
gse$Epithelium <- find_text("Epithelium")

# Collapse columns 
gse <- unite(gse,
             "SampleType",
             Blood:Tissue:Epithelium,
             na.rm = TRUE, 
             remove = TRUE)


###################################
# "Disease" as new column
###################################

# Emphysema
gse$Emphysema <- find_text("Emphysema")

# Bronchitis
gse$Bronchitis <- find_text("Bronchitis")

# COPD
gse$COPD <- find_text("COPD")

# Collapse columns 
gse <- unite(gse,
             "Disease",
             Emphysema:Bronchitis:COPD,
             na.rm = TRUE, 
             remove = TRUE)



###################################
# "Controls" as new column (either control or normal)
###################################
# control
gse$control <- find_text("control")

# normal
gse$normal <- find_text("normal")

# Collapse columns 
gse <- unite(gse,
             "Controls",
             control:normal,
             na.rm = TRUE, 
             remove = TRUE)


###################################
# "Patient" as new column
###################################
gse$Patient <- find_text("Patient")


###################################
# write table as "output_data/GSE_query_mined.txt"
###################################

write.table(gse,file = here::here("output_data/GSE_query_mined.txt"),quote = F,sep = "\t",row.names = F)


```


