---
title: "1-download_raw-data.RMD"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> Download raw data using GEOquery
</div>

### Background 

We are going to download raw data and meta information from [GEO](https://www.ncbi.nlm.nih.gov/geo/) using `GEOquery` package.


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>GSE list:</strong> GSE IDs from the experiments
</div>

## Analysis

### Setup

The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.

```{r setup}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("R/setup.R"))
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(GEOquery)
```


### GSE IDs

We need the GSE IDs to download raw files, the IDs are stored in a separate file named `GSE_IDs.txt`, we used `getGEOSuppFiles()` for download raw data.

```{r , class.source="bg-info",class.output="bg-info" }
tissue <- scan(DATA("GSE_IDs.txt"),
               character(), 
               sep = ",")

sapply(tissue,getGEOSuppFiles,baseDir = here::here("download"))
```


### STEP 2

## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Figure 1:</strong> Boxplot of samples  
  <strong>Figure 2:</strong> Heatmap of COPD patients 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





