---
title: "1-download_raw-data.RMD"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> The aim of this script
</div>

### Background 

Some relevant information with links [GEO](https://www.ncbi.nlm.nih.gov/geo/)


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Data 1:</strong> Matrix of normalized experiments  
  <strong>Data 2:</strong> Meta data 
</div>

## Analysis

### Setup

The script can be found in: `r getwd() `

```{r setup}
PATH = here::here()
DATA_DIR = file.path(PATH,"Data/")
OUTPUT_DIR = file.path(PATH,"output/")
FIG_DIR = file.path(PATH,"Fig")

knitr::opts_knit$set(root.dir = PATH)
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
```


### STEP 1

```{r , class.source="bg-info",class.output="bg-info" }
1+1
```


### STEP 2

## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Figure 1:</strong> Boxplot of samples  
  <strong>Figure 2:</strong> Heatmap of COPD patients 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





