---
title: "4-Differential Geneexpression analysis.RMD"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To perform differential expression analysis in COPD experiments, contransting COPD vs Control group
</div>

We previously had run the following scripts:

- 0: Data selection `/0-Data-selection.RMD`

- 1: Dowload RAW data `/1-download_raw-data.RMD`

- 2: Normalizing data `/2-normalizing-data.RMD`

- 3: Curating data `/3-curatig-data.RMD`


## Input

This script needs the following files, which were obtained by step 3 `/3-curatig-data.RMD`:

<div class="alert alert-warning">
  <strong>Table:</strong> GSE summary `2020-10-05_GSE_Summary.csv`   
  <strong>Data 2:</strong> ExpressionSet objects with normalized and curated data `2020-10-05_GSE_LungTissue-CURATED.RDS`    
</div>


## Setup

For running the script, type:

```{bash, eval = F}
ssh -X aaltamirano@dna.lavis.unam.mx
qrsh
cd /mnt/Genoma/amedina/DataDNA/R-projects/Meta-analysis_COPD/output_data/.out
 module load r/4.0.1 

## OR
ssh ana@10.200.0.42
cd /home/ana/DataDNA/R-projects/Meta-analysis_COPD/output_data/.out

nohup R -e "rmarkdown::render(here::here('vignettes/4-DE.RMD'))" > 4-DE.RMD.out &
```

The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.


```{r setup}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("R/setup.R"))
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(DESeq2)
library(limma)
library(oligo)
library(vsn)
```


### Experiments

We selected experiments from lung samples of COPD patients 
and that also have a control group to compare. These experiments are described 
in the following table:


```{r experiments}
gse_table <- read.csv(OUTPUT("2020-10-07_Step3_Summary.csv"), row.names = 1)
kable(gse_table, caption = "GSE information") %>%
  kableExtra::scroll_box(width = "100%", height = "100px")
```


## Local functions

### **DE**: Differential expression analysis

Using this funtion, we get a table with differential expression gene results using
`limma` package for fitting a linear model to get genes deferentially expressed
between a "Control" and a "COPD" group. 

<div class="alert alert-warning">
  <strong>Input:</strong> GSE ID, optional: *colCOPD* is the column name in which 
  the information of disease status can be found, *coeff* will show results of 
  contrast with coeffitient found in possiton 2   
  <strong>Output:</strong> Table of differential expression results with all
  genes 
</div>


```{r DE}
DE <- function(ExpressionSet,colCOPD="DISEASE",coeff= "DISEASECOPD"){
     # it creates the design matrix and performs limma
     fit <- lmFit(ExpressionSet, model.matrix(as.formula(paste("~ 1 +", colCOPD)),
                                              data = pData(ExpressionSet)))
     # eBayes in lmFit model
     ebf <- eBayes(fit)
     print(colnames(coef(fit)))
     # It gets the genes with the p-values
     volcanoplot(ebf,coef = coeff,highlight=20, pch=20)
     res <- topTable(ebf, number = Inf, p.value = 1, coef = coeff,confint=T)
     # It formats in a tibble
     res <- as_tibble(res,rownames="rownames")
}

```


For microarray experiments, all the following steps are the same, so we will 
wrap the following code into one function.

<div class="alert alert-warning">
  <strong>Input:</strong> Position of the GSE id inside of `geo` list (e.i. 1,2..)       
  <strong>Output:</strong> Table with DE results and a `.csv`
  genes 
</div>


```{r}

save_DE <- function(i){
  
g <- geo[[i]][[1]]
message("plotting data")
boxplot(g)
hist(g)

# DE using a local function. DISEASE is a column with disease status information
message("DE analysis")
de <- DE(g,colCOPD = "DISEASE")

# renaming colnames with the GSE id
message("Renaming colnames with GSE id")
colnames(de) <- str_c(colnames(de),"_",names(geo)[i])
colnames(de)

# saving the results in a csv
message("Saving data into csv")
write_csv(de,path= OUTPUT(c("DE/",TODAY,"_Step4_TableGenes_",i,"_",names(geo[i]),".csv")))

# saving the results to later combine all the information 
 return(de)
}

```

## Analysis

### DATA


```{r, class.source="bg-info",class.output="bg-info"}
geo <- readRDS(OUTPUT("2020-10-05_GSE_LungTissue-CURATED.RDS"))
names(geo)
```


## Differential expression analysis

Using `DE()` function (described above), we performed a lineal regression model 
to calculate the logarithm fold change of all genes between a "Control" and a
"COPD" group. We also rename colnames adding the GSE ID at the end and finally, 
we save the output in a `.CSV` file. 


### `r names(geo)[1] `

```{r}
i <- 1
de1 <- save_DE(i)
```


### `r names(geo)[2] `

```{r}
i <- 2
de1 <- save_DE(i)
```


### `r names(geo)[3] `

```{r}
i <- 3
de3 <- save_DE(i)
```



### `r names(geo)[4] `

This experiment is RNAseq, the data will be download from Recount2. I’m following Recount2 vignette
Using `DESeq2`package we calculated DEG.  

```{r}
## Specify design and switch to DESeq2 format
i <- 4
rse <-  geo[[i]][[1]]

 dds <- DESeqDataSet(rse, ~ DISEASE)
 
 ## Perform DE analysis
 dds <- DESeq(dds)
 boxplot(log10(assays(dds)[["cooks"]]), range=0, las=2)
 
 ntd <- normTransform(dds)
 meanSdPlot(assay(ntd))
 
 res <- results(dds)
 plotMA(res, ylim=c(-2,2))
 
 # Calculates de CI
 res$error <- qnorm(0.975)*res$lfcSE
 res$CI.L <- res$log2FoldChange-res$error
 res$CI.R <- res$log2FoldChange+res$error
 
 res
 
 
r <- as_tibble(res, rownames="rownames")
r <- full_join(r,as.data.frame(rowData(geo[[i]][[1]])), by=c("rownames"="gene_id")) 
r$logFC <- r$log2FoldChange

colnames(r) <- str_c(colnames(r),"_",names(geo[i]))
colnames(r)
r <- r[,-12]
colnames(r)

write_csv(r,path= OUTPUT(c("DE/",TODAY,"_Step4_TableGenes_",i,"_",names(geo[i]),".csv")))


de4 <- r
```



### `r names(geo)[5] `

```{r}
i <- 5
de5 <- save_DE(i)
```


### `r names(geo)[6] `

```{r}
i <- 6
de6 <- save_DE(i)
```

### `r names(geo)[7] `

```{r}
i <- 7
de7 <- save_DE(i)
```

### `r names(geo)[8] `

```{r}
i <- 8
de8 <- save_DE(i)
```

### `r names(geo)[9] `

```{r}
i <- 9
de9 <- save_DE(i)
```



### `r names(geo)[10] `

```{r}
i <- 10
de10 <- save_DE(i)
```



### `r names(geo)[11] `

```{r}
i <- 11
de11 <- save_DE(i)
```



### `r names(geo)[12] `

```{r }
i <- 12
rse <-  geo[[i]][[1]]

counts <- t(apply(exprs(rse), 1,as.numeric))
colnames(counts) <- colnames(rse)
colData <- pData(rse)

rse <- SummarizedExperiment(assays=list(counts=counts), colData=colData)
rowData(rse) <- fData(geo[[i]][[1]])

 dds <- DESeqDataSet(rse, ~ DISEASE)
 
 ## Perform DE analysis
 dds <- DESeq(dds)
 boxplot(log10(assays(dds)[["cooks"]]), range=0, las=2)
 
 ntd <- normTransform(dds)
 meanSdPlot(assay(ntd))
 
 res <- results(dds)
 plotMA(res, ylim=c(-2,2))
 
 # Calculates de CI
 res$error <- qnorm(0.975)*res$lfcSE
 res$CI.L <- res$log2FoldChange-res$error
 res$CI.R <- res$log2FoldChange+res$error
 
 res

r <- as_tibble(res, rownames="rownames")
r <- full_join(r,as.data.frame(rowData(rse)), by=c("rownames"="GENEID")) 
r$logFC <- r$log2FoldChange

colnames(r) <- str_c(colnames(r),"_",names(geo[i]))
colnames(r)


write_csv(r,path= OUTPUT(c("DE/",TODAY,"_Step4_TableGenes_",i,"_",names(geo[i]),".csv")))

# saving the results to later combine all the information 
de12 <- r
```









## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Tables with DE results:</strong> Tables with log fold change and p-values 
  calculated per experiment and saved in `csv` files   
</div>


## Session Info

```{r sessionInfo}
sessionInfo()
```





