---
title: "6-meta-analysis.RMD"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To do a meta-analysis for COPD experiments
</div>


We previously had run the following scripts:

- 0: Data selection `/0-Data-selection.RMD`

- 1: Dowload RAW data `/1-download_raw-data.RMD`

- 2: Normalizing data `/2-normalizing-data.RMD`

- 3: Curating data `/3-curatig-data.RMD`

- 4: Differential gene expression analysis  `/4-DE.RMD`

- 5: Merging data `/5-merge-data.RMD`


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Table:</strong> Table with all DE results. Output from step 5   
</div>

## Analysis

### Setup

For running the script, type:

```{bash, eval = F}
ssh -X aaltamirano@dna.lavis.unam.mx
qrsh
cd /mnt/Genoma/amedina/DataDNA/R-projects/Meta-analysis_COPD/output_data/.out
 module load r/4.0.1 

## OR
ssh ana@10.200.0.42
cd /home/ana/DataDNA/R-projects/Meta-analysis_COPD/output_data/.out

nohup R -e "rmarkdown::render(here::here('vignettes/6-meta-analysis.RMD'))" > 6-meta-analysis.RMD.out &
```

The script can be found in: `r getwd() `. And the directories were set using
`R/setup.R` (e.i. DATA, OUPUT, DOWNLOAD) as a small functions that will paste the names into complete paths.


```{r setup}
knitr::opts_knit$set(root.dir = here::here())
source(here::here("R/setup.R"))
```

And this analysis is run in: `r getwd() `


### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(ComplexHeatmap)
library(metafor)
library(pheatmap)
library(RColorBrewer)
```


### Read data

```{r , class.source="bg-info",class.output="bg-info" }
full_tables <- read_csv(OUTPUT("2020-10-08_Step5_Full_Tables.csv"))
colnames(full_tables)

```



### p-Values

We first check pvalues histograms. [Ref](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6164648/)

```{r , class.source="bg-info",class.output="bg-info", fig.width= 8 }
FC <- str_subset(colnames(full_tables),"logFC")

# Genes with a adjusted p value
adjp <- full_tables %>%
  filter(GENE.SYMBOL != "NA") %>%
  gather("GSE"  ,"adjP",-GENE.SYMBOL) %>%
  filter(GSE %in% str_subset(colnames(full_tables),"adj.P.Val|padj")) %>%
  mutate(GSE=gsub(".*_","",GSE))
  
adjp


#pdf("Pvalues.pdf",width = 8,height = 6)
adjp %>%    
  group_by(GSE) %>%
  ggplot( aes(x=adjP, fill=GSE)) +
  geom_histogram( color="#e9ecef", alpha=0.6) + #position = 'identity'
  geom_vline(xintercept=0.5, linetype="dashed", color = "red") +
  facet_wrap(~GSE, scales = "free_y")
#dev.off()

adj_pvalue <- 0.5
adjp <- filter(adjp,adjP < adj_pvalue)
unique(adjp$GSE)
adjp

```


Then, we look for how many genes are up regulated with a p-value less than `r  adj_pvalue`  in the 

```{r fig.width=10 }

# log fold change
logFC <- full_tables %>%
  filter(GENE.SYMBOL != "NA") %>%
  gather("GSE"  ,"logFC",-GENE.SYMBOL) %>%
  filter(GSE %in% str_subset(colnames(full_tables),"logFC")) %>% 
  mutate(GSE=gsub(".*_","",GSE))

logFC

# recovering logFC values from genes with p value less than 0.01 

ap <- left_join(adjp,logFC,by = c("GENE.SYMBOL" = "GENE.SYMBOL", "GSE" = "GSE")) %>%
  select(-adjP) %>%
  spread(GSE,logFC) %>% 
  replace(is.na(.), 0)

ap

# selecting up regulated genes
up <- ap %>%
  mutate_at(vars(-GENE.SYMBOL),list(~if_else( .>0, 1, 0)))
  
down <- ap %>%
  mutate_at(vars(-GENE.SYMBOL),list(~if_else( .<0, 1, 0)))

### UP 
m <- make_comb_mat(up,mode = "distinct")
m <- m[comb_degree(m) >= 5 | comb_degree(m) == 1]

#pdf("upsetUP.pdf",width = 14,height = 6)
ht <- draw(UpSet(m,
      pt_size = unit(3, "mm"), 
      lwd = 3,
      comb_col = c("red", "grey","grey","grey", "black","blue")[comb_degree(m)])
      #order.by = "freq"
    )

od = column_order(ht)
cs = comb_size(m)
decorate_annotation("Intersection\nsize", {
    grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(2, "pt"), 
        default.units = "native", just = "bottom", gp = gpar(fontsize = 8))
})

#dev.off()


### DOWN
m <- make_comb_mat(down,mode = "distinct")
m <- m[comb_degree(m) >= 5 | comb_degree(m) == 1] 

#pdf("upsetDOWN.pdf",width = 14,height = 6)
ht <- draw(UpSet(m,
      pt_size = unit(3, "mm"), 
      lwd = 3,
      comb_col = c("red", "grey","grey","grey", "black","blue")[comb_degree(m)])
      #order.by = "freq"
    )
#dev.off()

od = column_order(ht)
cs = comb_size(m)
decorate_annotation("Intersection\nsize", {
    grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(2, "pt"), 
        default.units = "native", just = "bottom", gp = gpar(fontsize = 8))
})

```


## Meta-analysis

We first calculate the Standar error using confidence intervals

```{r}

CI.R <- str_subset(colnames(full_tables),"CI.R")
CI.L <- str_subset(colnames(full_tables),"CI.L")

for(i in 1:length(CI.R)) {
  full_tables[str_c("SE_",gsub(".*_","",CI.R))] <- (full_tables[,CI.R]-full_tables[,CI.L]) /(2*1.96)
}
```

Meta-analysis for a single gene


```{r}

FC <- str_subset(colnames(full_tables),"logFC")
SE <- str_subset(colnames(full_tables),"SE_")
GSE <- gsub(".*_","",FC)

gs <- "CXCL12"

## Funtion to perform a meta-analysis per gene
meta.plot <- function(gs){
gene <-  full_tables[full_tables$GENE.SYMBOL == gs,]
#gene <-  na.omit(full_tables[full_tables$GENE.SYMBOL == gs,])
gene <- gene[!is.na(gene$GENE.SYMBOL),]

yi <- as.numeric(gene[,FC])
sei <- as.numeric(gene[,SE])

# meta analysis in the simplest way
res <- metafor::rma(yi, sei=sei)
funnel(res)
forest(res,slab = GSE, main=unique(gene$GENE.SYMBOL))
}

meta.plot(gs)
```

Meta-analysis for all genes. We are saving the results in a new dataframe.

```{r}

#gs <- unique(na.omit(full_tables)$GENE.SYMBOL)
gs <- full_tables$GENE.SYMBOL


meta <- data.frame(genes = 0,
                   QEp=0,
                   pval=0,
                   tau2=0,
                   I2=0,
                   num_exp=0)

gm <- full_tables$GENE.SYMBOL
full_tables[full_tables == 0] <- NA

length(gm)

for (i in gm) {
  gene <- full_tables[which(full_tables$GENE.SYMBOL == i),]
  yi <- as.numeric(gene[,FC])
  sei <- as.numeric(gene[,SE])
  num_exp <- sum(!is.na(yi))
  # meta analysis with random models "REML"
  res <- suppressWarnings(metafor::rma(yi, sei=sei,control=list(stepadj=.5,maxiter=1000)))
  num_exp <- sum(!is.na(yi))
  meta <- rbind(meta,c(i,res·QEp,res$zval,res$pval,res$se,res$H2,res$tau2,res$I2,num_exp))
}



meta <- meta[-1,]
meta$QEp <- as.numeric(meta$QEp)
meta$pval <- as.numeric(meta$pval)

# FDR correction
meta$qval <- p.adjust(meta$pval,method="fdr")


write.csv(meta, OUTPUT(c(TODAY,"_Step6_meta-analysis.csv")))
# meta <- read.csv(OUTPUT("2020-10-08_Step6_meta-analysis.csv"))

hist(meta$qval, main="q-value from Cochrane's test", xlab="FDR(P-value)")
hist(meta$QEp, main="Test for Heterogeneity: Q test", xlab="QEp")

qval_cutoff <- 0.01
genesqval <- meta[meta$qval <= qval_cutoff,]
#genesqval <- genesqval[genesqval$QEp < 0.2,]
genesqval <-  genesqval[order(genesqval$qval),]

kable(genesqval, caption = "Meta-analysis")

up <- "IL6"
down <- "PRKCZ"
ace2 <- "ACE2"

meta.plot(up)
meta.plot(down)
meta.plot(ace2)

topgenes <- full_tables[full_tables$GENE.SYMBOL %in% genesqval$genes,c("GENE.SYMBOL",FC)]
topgenes <- column_to_rownames(topgenes,"GENE.SYMBOL")

topgenes <- scale(topgenes)

# keeping genes with less than 2 NAs
topgenes <- topgenes[which(rowSums(is.na(topgenes)) <= 2),]
#topgenes <- na.omit(topgenes)

#ordering rows
#topgenes <- topgenes[order(rowMeans(topgenes),decreasing = T),]

#"BrBG"
color <- rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(15))
breaks <- c(seq(min(topgenes,na.rm = T),-0.1,length.out = 7),0,seq(0.1,max(topgenes,na.rm = T),length.out = 7))

```

```{r fig.height=15}

pheatmap(topgenes,
         main = str_c("Genes with q-val less than ",qval_cutoff),
         color = color,
         breaks = breaks,
         fontsize_row = 7,
         fontsize_col = 10,
         cluster_rows = T,
         clustering_distance_rows = "euclidean",
         clustering_method = "ward.D",
         cluster_cols = T,
         cutree_rows = 3,
         show_colnames = T,
         border_color = NA)

```

### TFs

We also checked for expression in TFs 

````{r}
tfs <- read.csv(DATA("hs_hgnc_curated_tfs.txt"),col.names = "TF")

qval_cutoff <- 0.1
genesqval <- meta[meta$qval <= qval_cutoff,]
tf_copd <- genesqval$genes[which(genesqval$genes %in% tfs$TF)]

topgenes <- full_tables[full_tables$GENE.SYMBOL %in% tf_copd,c("GENE.SYMBOL",FC)]
topgenes <- column_to_rownames(topgenes,"GENE.SYMBOL")
topgenes <- scale(topgenes)
#topgenes <- na.omit(topgenes)


color <- rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(15))
breaks <- c(seq(min(topgenes,na.rm = T),-0.1,length.out = 7),0,seq(0.1,max(topgenes,na.rm = T),length.out = 7))

pheatmap(topgenes,
         main = str_c("TFs with q-val less than ",qval_cutoff),
         color = color,
         breaks = breaks,
         fontsize_row = 8,
         fontsize_col = 10,
         cluster_rows = T,
         clustering_distance_rows = "euclidean",
         clustering_method = "ward.D",
         cluster_cols = T,
         cutree_rows = 2,
         show_colnames = T,
         border_color = NA)


```



## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Plots:</strong> Upset plots, forest plot, heatmap, etc.     
  <strong>Table:</strong> Results from the meta-analysis `r OUTPUT(c(TODAY,"_Step6_meta-analysis.csv"))` 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





