---
title: "Enrichment Analysis"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> The aim is to do an enrichment analysis using results from  
  the meta-analysis.
</div>

### Background 

This analysis is based on [hyperR BioC 2020 workshop](https://montilab.github.io/hypeR-workshop/articles/hyper_workshop.html).

## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Data 1:</strong> Matrix of normalized experiments  
  <strong>Data 2:</strong> Meta data 
</div>

## Analysis

### Setup

The script can be found in: `r getwd() `

```{r setup}
PATH = here::here() # "."
DATA_DIR = file.path(PATH,"Data/")
OUTPUT_DIR = file.path(PATH,"output/")
FIG_DIR = file.path(PATH,"Fig")

knitr::opts_knit$set(root.dir = PATH)
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
# BiocManager::install("hypeR")
library(hypeR)
library(WGCNA)
library(tidyverse)
library(reactable)
```


### STEP 1

```{r , class.source="bg-info",class.output="bg-info" }
meta <- read.csv("output_data/meta-analysis_preprocessData-2020-08-13.csv")[,-1]
signature <- meta[meta$qval <= 0.5,"QEp"]
names(signature) <- meta[meta$qval <= 0.5,"genes"]

head(signature)
```

### STEP 2

We are importing the latest genesets from Kegg.

```{r}
# KEGG
KEGG <- msigdb_gsets("Homo sapiens", "C2", "CP:KEGG")
print(KEGG)
#HALLMARL
HALLMARK <- msigdb_gsets(species="Homo sapiens", category="H",clean = T)
print(HALLMARK)

#ENRICHR
available <- enrichr_available()
reactable(available)
enrichR <- enrichr_gsets("TRANSFAC_and_JASPAR_PWMs")
enrichR<- enrichr_gsets("GeneSigDB")
enrichR<- enrichr_gsets(available[67,1])
enrichR<- enrichr_gsets(available[68,1])
enrichR<- enrichr_gsets(available[69,1])
enrichR<- enrichr_gsets(available[71,1])
enrichR<- enrichr_gsets(available[103,1])
enrichR<- enrichr_gsets(available[79,1])
enrichR<- enrichr_gsets(available[80,1])
enrichR<- enrichr_gsets(available[102,1])
enrichR<- enrichr_gsets(available[107,1])

enrichR<- enrichr_gsets(available[1,1])
enrichR<- enrichr_gsets(available[130,1])
enrichR<- enrichr_gsets(available[110,1])
enrichR<- enrichr_gsets(available[132,1])
```


```{r}
hyp_obj <- hypeR(names(signature),enrichR, test="hypergeometric", background=nrow(meta), fdr=0.5, plotting=TRUE)
hyp_obj$plots[[1]]
print(hyp_obj)
```

```{r}
hyp_obj <- hypeR(signature, enrichR, test="kstest", fdr=0.5, plotting=TRUE)
hyp_obj$plots[[1]]
print(hyp_obj)
```

```{r}
# Show interactive table
hyp_show(hyp_obj)

# Plot dots plot
hyp_dots(hyp_obj)

# Plot enrichment map
hyp_emap(hyp_obj)

# Plot hiearchy map
hyp_hmap(hyp_obj, top = 30)

# Save to excel
hyp_to_excel(hyp_obj)

# Save to table
hyp_to_table(hyp_obj)

# Generate markdown report
#hyp_to_rmd(hyp_obj,"vignettes/hypr-hallmark.rmd")
```



## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Figure 1:</strong> Boxplot of samples  
  <strong>Figure 2:</strong> Heatmap of COPD patients 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





