---
title: "Gene Annotation"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To join results from differential gene expression
  analysis into a single dataframe using gene symbol annotation.
</div>

### Background 

In a previous vignette `RMA_DE-23-June-2020.Rmd`, we perfomed differential gene 
expression analysis per experiment and calculated log fold-change, p-values,
adjusted p-values and confidence intervals per gene.

Here, we want to merge the results into one table using gene symbol annotation. 
Each platform has different gene ids that are not comparable so we will use 
gene symbol annotation (i.e. MMP8, AHR) as reference to join all results.


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Data 1:7:</strong> Matrix of differential gene expression results.  
</div>

## Process

### Setup

All the data has already been downloaded in the cluster *10.200.0.42*, for 
accessing:

```{bash, eval=F}
ssh ana@10.200.0.42

cd /home/ana/R-projects/Meta-analysis_COPD
```


For running the script, type:

```{bash, eval= F}
nohup R -e "rmarkdown::render('vignettes/GeneSymbol-Annotation_9-July-2020.Rmd')" & > genesymbol.out
```


The script can be found in: `r getwd() `

```{r setup}
setwd("..")
PATH = getwd()
DATA_DIR = file.path(PATH,"data")
OUTPUT_DIR = file.path(PATH,"output_data")
FIG_DIR = file.path(PATH,"fig")
TODAY = Sys.Date()

knitr::opts_knit$set(root.dir = PATH)
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
```

### Experiments

We selected 7 experiments that are in PulmonDB and are lung samples from COPD patients 
and that also have a control group to compare. These experiments are described 
in the following table:


```{r experiments}
gse_table <- read.csv(file.path(DATA_DIR,"GSE_table.csv"), row.names = 1)
kable(gse_table, caption = "GSE information")
```


### STEP 1

We read all Tables located in `r OUTPUT_DIR`.

```{r readfiles}
temp = list.files(OUTPUT_DIR,pattern="TableGenes*",full.names =T)

data <- temp %>% map(read_csv)

```

### STEP 2

All Tables need to have `Gene.Symbol` column, for some experiments we will need 
to parse it. For expample: Gene.Symbol from `gene_assigment_GSE27597`.

```{r}
colnames(data[[3]])
data[[3]][1,"gene_assignment_GSE27597"]

# Parse column to have Gene.Symbol
gene_name <- dplyr::mutate(data[[3]],
                         geneInfo = strsplit(gene_assignment_GSE27597, " /// "),
                         gene1 = sapply(geneInfo, `[`, 1),
                         Gene.Symbol_GSE27597 = sapply(strsplit(gene1, " // "), `[`, 2)
                         )
gene_name <- dplyr::select(gene_name, rownames_GSE27597,Gene.Symbol_GSE27597)

# Rename dataframe
data[[3]] <- as_tibble(merge(gene_name,
                     data[[3]],
                     by="rownames_GSE27597"))

```


### STEP 3

Using `column_selection` function, we can subset columns that has Gene symbol,
logFC, adjsted p-value, and confidence intervals with unique id. 

```{r column_selection, class.source="bg-info",class.output="bg-info" }

column_selection <- function(df){
  # Compete column name (i.e "Gene.Symbol_GSE1122")
  GeneS <- str_subset(colnames(df),regex("Symbol",ignore_case  =  TRUE))
  logFC <- str_subset(colnames(df),"logFC|log2F")
  adjP <- str_subset(colnames(df),"adj.P.Val|padj")
  CI.L <- str_subset(colnames(df),"CI.L")
  CI.R <- str_subset(colnames(df),"CI.R")
  
  selectedColumns <- c(GeneS,logFC,adjP,CI.L,CI.R)
  #Return subset of dataframe to only keep the first duplicate
  DF <- df[!duplicated(df[,GeneS]),selectedColumns] %>%
    rename(Gene.Symbol=GeneS)
  
  return(DF)
}

```

### STEP 4

Join all data into one dataset!

```{r merge}
full_tables <- data %>% 
  map(column_selection)%>% 
  purrr::reduce(full_join, by = "Gene.Symbol")

full_tables  

#left_tables <- data %>% 
#  map(column_selection)%>% 
#  purrr::reduce(left_join, by = "Gene.Symbol")


write_csv(full_tables,
          path=str_c(OUTPUT_DIR,"/Full_Tables_",TODAY,".csv"))
```


## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Dataframe:</strong> A dataframe with logFC, adjP,CI.R and CI.L from all experiments
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





