---
title: "Meta-analysis"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To perform a Meta-analysis using a random effect model 
</div>

### Background 

Some relevant information with links [GEO](https://www.ncbi.nlm.nih.gov/geo/)


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Dataframe:</strong> Matrix of differential gene expression analysis
</div>

## Analysis

### Setup

All the data has already been downloaded in the cluster *10.200.0.42*, for 
accessing:

```{bash, eval=F}
ssh ana@10.200.0.42

cd /home/ana/R-projects/Meta-analysis_COPD
```


For running the script, type:

```{bash, eval= F}
nohup R -e "rmarkdown::render('vignettes/Meta-analysis_Lung_Rawdata_9-July-2020.Rmd')" & > meta-analysis.out
```


The script can be found in: `r getwd() `

```{r setup}
PATH = here::here()
DATA_DIR = file.path(PATH,"data")
OUTPUT_DIR = file.path(PATH,"output_data")
FIG_DIR = file.path(PATH,"fig")
TODAY = Sys.Date()

knitr::opts_knit$set(root.dir = PATH)
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(ComplexHeatmap)
library(metafor)
library(pheatmap)
library(RColorBrewer)
```

### Experiments

We selected 7 experiments that are in PulmonDB and are lung samples from COPD patients 
and that also have a control group to compare. These experiments are described 
in the following table:


```{r experiments}
gse_table <- read.csv(file.path(DATA_DIR,"GSE_table.csv"), row.names = 1)
kable(gse_table, caption = "GSE information")
```


### STEP 1

We will import DE results from all experiments. 

```{r , class.source="bg-info",class.output="bg-info" }
full_tables <- read_csv(str_c(OUTPUT_DIR,"/Full_Tables_2020-07-13.csv"))
colnames(full_tables)
colnames(full_tables)[colnames(full_tables) == 'log2FoldChange_GSE57148'] <- 'logFC_GSE57148'
full_tables
```


### STEP 2

We first check pvalues histograms. [Ref](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6164648/)

```{r , class.source="bg-info",class.output="bg-info" }
FC <- str_subset(colnames(full_tables),"logFC")

# Genes with a adjusted p value
adjp <- full_tables %>%
  filter(Gene.Symbol != "NA") %>%
  gather("GSE"  ,"adjP",-Gene.Symbol) %>%
  filter(GSE %in% str_subset(colnames(full_tables),"adj.P.Val|padj")) %>%
  mutate(GSE=gsub(".*_","",GSE))
  
adjp


#pdf("Pvalues.pdf",width = 8,height = 6)
adjp %>%    
  group_by(GSE) %>%
  ggplot( aes(x=adjP, fill=GSE)) +
  geom_histogram( color="#e9ecef", alpha=0.6) + #position = 'identity'
  geom_vline(xintercept=0.5, linetype="dashed", color = "red") +
  facet_wrap(~GSE, scales = "free_y")
#dev.off()

adj_pvalue <- 0.5
adjp <- filter(adjp,adjP < adj_pvalue)
adjp
```


Then, we look for how many genes are up regulated with a p-value less than `r  adj_pvalue`  in the 

```{r  }

# log fold change
logFC <- full_tables %>%
  filter(Gene.Symbol != "NA") %>%
  gather("GSE"  ,"logFC",-Gene.Symbol) %>%
  filter(GSE %in% str_subset(colnames(full_tables),"logFC")) %>% 
  mutate(GSE=gsub(".*_","",GSE))

logFC

# recovering logFC values from genes with p value less than 0.01 

ap <- left_join(adjp,logFC,by = c("Gene.Symbol" = "Gene.Symbol", "GSE" = "GSE")) %>%
  select(-adjP) %>%
  spread(GSE,logFC) %>% 
  replace(is.na(.), 0)

ap

# selecting up regulated genes
up <- ap %>%
  mutate_at(vars(-Gene.Symbol),funs(if_else( .>0, 1, 0)))
  
down <- ap %>%
  mutate_at(vars(-Gene.Symbol),funs(if_else( .<0, 1, 0)))

### UP 
m <- make_comb_mat(up,mode = "distinct")
m <- m[comb_degree(m) >= 4 | comb_degree(m) == 1]

#pdf("upsetUP.pdf",width = 14,height = 6)
ht <- draw(UpSet(m,
      pt_size = unit(3, "mm"), 
      lwd = 3,
      comb_col = c("red", "grey","grey","grey", "black","blue")[comb_degree(m)])
      #order.by = "freq"
    )
od = column_order(ht)
cs = comb_size(m)
decorate_annotation("Intersection\nsize", {
    grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(2, "pt"), 
        default.units = "native", just = "bottom", gp = gpar(fontsize = 8))
})

#dev.off()


### DOWN
m <- make_comb_mat(down,mode = "distinct")
m <- m[comb_degree(m) >= 4 | comb_degree(m) == 1] 

#pdf("upsetDOWN.pdf",width = 14,height = 6)
ht <- draw(UpSet(m,
      pt_size = unit(3, "mm"), 
      lwd = 3,
      comb_col = c("red", "grey","grey","grey", "black","blue")[comb_degree(m)])
      #order.by = "freq"
    )
#dev.off()

od = column_order(ht)
cs = comb_size(m)
decorate_annotation("Intersection\nsize", {
    grid.text(cs[od], x = seq_along(cs), y = unit(cs[od], "native") + unit(2, "pt"), 
        default.units = "native", just = "bottom", gp = gpar(fontsize = 8))
})

```


## Meta-analysis

We first calculate the Standar error using confidence intervals

```{r}

CI.R <- str_subset(colnames(full_tables),"CI.R")
CI.L <- str_subset(colnames(full_tables),"CI.L")

for(i in 1:length(CI.R)) {
  full_tables[str_c("SE_",gsub(".*_","",CI.R))] <- (full_tables[,CI.R]-full_tables[,CI.L]) /(2*1.96)
}
```

Meta-analysis for a single gene


```{r}

FC <- str_subset(colnames(full_tables),"logFC")
SE <- str_subset(colnames(full_tables),"SE_")
GSE <- gsub(".*_","",FC)

gs <- "CXCL12"
meta.plot <- function(gs){
gene <-  full_tables[full_tables$Gene.Symbol == gs,]
#gene <-  na.omit(full_tables[full_tables$Gene.Symbol == gs,])
gene <- gene[!is.na(gene$Gene.Symbol),]

yi <- as.numeric(gene[,FC])
sei <- as.numeric(gene[,SE])

# meta analysis in the simplest way
res <- metafor::rma(yi, sei=sei)
funnel(res)
forest(res,slab = GSE, main=unique(gene$Gene.Symbol))
}

meta.plot(gs)
```

Meta-analysis for all genes

```{r}

#gs <- unique(na.omit(full_tables)$Gene.Symbol)
gs <- unique(full_tables$Gene.Symbol)


meta <- data.frame(genes = 0,
                   QEp=0,
                   pval=0,
                   tau2=0,
                   I2=0,
                   num_exp=0)

fc <- full_tables[,FC]
gm<- full_tables$Gene.Symbol[which(rowSums(!is.na(fc)) > 2)]
gm <- na.omit(gm)

length(gm)

for (i in gm) {
  gene <- full_tables[which(full_tables$Gene.Symbol == i),]
  yi <- as.numeric(gene[,FC])
  sei <- as.numeric(gene[,SE])
  num_exp <- sum(!is.na(yi))
  # meta analysis with random models "REML"
  res <- metafor::rma(yi, sei=sei)
  num_exp <- sum(!is.na(yi))
  meta <- rbind(meta,c(i,res$QEp,res$pval,res$tau2,res$I2,num_exp))
}



meta <- meta[-1,]
meta$QEp <- as.numeric(meta$QEp)
meta$pval <- as.numeric(meta$pval)

# FDR correction
meta$qval <- p.adjust(meta$pval,method="fdr")


write.csv(meta,str_c(OUTPUT_DIR,"/meta-analysis_preprocessData-",TODAY,".csv"))

#meta <- read.csv(str_c(OUTPUT_DIR,"/meta-analysis_preprocessData-2020-08-13.csv"))

hist(meta$qval, main="q-value from Cochrane's test", xlab="FDR(P-value)")
hist(meta$QEp, main="Test for Heterogeneity: Q test", xlab="QEp")

genesqval <- meta[meta$qval <= 0.01,]
#genesqval <- genesqval[genesqval$QEp < 0.2,]
genesqval <-  genesqval[order(genesqval$qval),]

kable(genesqval, caption = "Meta-analysis")


for (i in 1:3){
meta.plot(genesqval$genes[i])
}

color <- rev(colorRampPalette(brewer.pal(n = 11, name = "RdBu"))(15))
breaks <- c(seq(min(topgenes),-0.1,length.out = 7),0,seq(0.1,max(topgenes),length.out = 7))
#"BrBG"
topgenes <- full_tables[full_tables$Gene.Symbol %in% genesqval$genes,c("Gene.Symbol",FC)]
topgenes <- column_to_rownames(topgenes,"Gene.Symbol")
topgenes <- na.omit(topgenes)
#ordering rows
#topgenes <- topgenes[order(rowMeans(topgenes),decreasing = T),]
```

```{r fig.height=15}

pheatmap(na.omit(topgenes),
         color = color,
         breaks = breaks,
         fontsize_row = 7,
         fontsize_col = 10,
         cluster_rows = T,
         clustering_distance_rows = "euclidean",
         clustering_method = "ward.D",
         cluster_cols = T,
         cutree_rows = 2,
         show_colnames = T,
         border_color = NA)

```


## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Figure 1:</strong> Boxplot of samples  
  <strong>Figure 2:</strong> Heatmap of COPD patients 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





