---
title: "Meta-analysis"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To perform a Meta-analysis using a random effect model 
</div>

### Background 

Some relevant information with links [GEO](https://www.ncbi.nlm.nih.gov/geo/)


## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Dataframe:</strong> Matrix of differential gene expression analysis
</div>

## Analysis

### Setup

All the data has already been downloaded in the cluster *10.200.0.42*, for 
accessing:

```{bash, eval=F}
ssh ana@10.200.0.42

cd /home/ana/R-projects/Meta-analysis_COPD
```


For running the script, type:

```{bash, eval= F}
nohup R -e "rmarkdown::render('vignettes/Meta-analysis_Lung_Rawdata_9-July-2020.Rmd')" & > meta-analysis.out
```


The script can be found in: `r getwd() `

```{r setup}
setwd("..")
PATH = getwd()
DATA_DIR = file.path(PATH,"data")
OUTPUT_DIR = file.path(PATH,"output_data")
FIG_DIR = file.path(PATH,"fig")
TODAY = Sys.Date()

knitr::opts_knit$set(root.dir = PATH)
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(ComplexHeatmap)
```

### Experiments

We selected 7 experiments that are in PulmonDB and are lung samples from COPD patients 
and that also have a control group to compare. These experiments are described 
in the following table:


```{r experiments}
gse_table <- read.csv(file.path(DATA_DIR,"GSE_table.csv"), row.names = 1)
kable(gse_table, caption = "GSE information")
```


### STEP 1

We will import DE results from all experiments. 

```{r , class.source="bg-info",class.output="bg-info" }
full_tables <- read_csv(str_c(OUTPUT_DIR,"/Full_Tables_2020-07-13.csv"))
colnames(full_tables)
colnames(full_tables)[colnames(full_tables) == 'log2FoldChange_GSE57148'] <- 'logFC_GSE57148'
full_tables
```


### STEP 2

We first check pvalues histograms. 

```{r , class.source="bg-info",class.output="bg-info" }
FC <- str_subset(colnames(full_tables),"logFC")

# Genes with a adjusted p value


adjp <- full_tables %>%
  filter(Gene.Symbol != "NA") %>%
  gather("GSE"  ,"adjP",-Gene.Symbol) %>%
  filter(GSE %in% str_subset(colnames(full_tables),"adj.P.Val|padj")) %>%
  mutate(GSE=gsub(".*_","",GSE))
  
adjp


#pdf("Pvalues.pdf",width = 8,height = 6)
adjp %>%    
  group_by(GSE) %>%
  ggplot( aes(x=adjP, fill=GSE)) +
  geom_histogram( color="#e9ecef", alpha=0.6) + #position = 'identity'
  geom_vline(xintercept=0.5, linetype="dashed", color = "red") +
  facet_wrap(~GSE, scales = "free_y")
#dev.off()

adj_pvalue <- 0.01
adjp <- filter(adjp,adjP < adj_pvalue)
adjp
```


Then, we look for how many genes are up regulated with a p-value less than `r  adj_pvalue`  in the 

```{r  }

# log fold change
logFC <- full_tables %>%
  filter(Gene.Symbol != "NA") %>%
  gather("GSE"  ,"logFC",-Gene.Symbol) %>%
  filter(GSE %in% str_subset(colnames(full_tables),"logFC")) %>% 
  mutate(GSE=gsub(".*_","",GSE))

logFC

# recovering logFC values from genes with p value less than 0.01 

ap0.01 <- left_join(adjp,logFC,by = c("Gene.Symbol" = "Gene.Symbol", "GSE" = "GSE")) %>%
  select(-adjP) %>%
  spread(GSE,logFC) %>% 
  replace(is.na(.), 0)

ap0.01

# selecting up regulated genes
up <- ap0.01 %>%
  mutate_at(vars(-Gene.Symbol),funs(if_else( .>0, 1, 0)))
  
down <- ap0.01 %>%
  mutate_at(vars(-Gene.Symbol),funs(if_else( .<0, 1, 0)))

### UP 
m <- make_comb_mat(up,mode = "intersect")
m <- m[comb_degree(m) >= 1]

#pdf("upsetUP.pdf",width = 14,height = 6)
UpSet(m,
      pt_size = unit(5, "mm"), 
      lwd = 4,
      comb_col = c("red", "grey","grey","grey", "black","blue")[comb_degree(m)]
      #order.by = "freq"
    )
#dev.off()


### DOWN
m <- make_comb_mat(down,mode = "intersect")
m <- m[comb_degree(m) >= 1]

#pdf("upsetDOWN.pdf",width = 14,height = 6)
UpSet(m,
      pt_size = unit(5, "mm"), 
      lwd = 4,
      comb_col = c("red", "grey","grey","grey", "black","blue")[comb_degree(m)]
      #order.by = "freq"
    )
#dev.off()


```


## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Figure 1:</strong> Boxplot of samples  
  <strong>Figure 2:</strong> Heatmap of COPD patients 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





