---
title: "Pre-processing and differential gene expression analysis of RAW data"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> To pre-process raw data from *GEO* using RMA and to 
  perform a differential gene expression analysis comparing COPD vs CONTROL group
</div>


### COPD

Chronic Obstructive Pulmonary Disease (COPD) is characterized by emphysema and 
chronic bronchitis, it's diagnosed using spirometry and clinical information
which lead to a heterogeneous COPD patients. The ranking of non commutable 
diseases from the WHO estimates that COPD is in the top of mortality causes and 
tobacco is the main risk factor but different genetic variants have been
associated with this disease. 

Different researchers have analyzed COPD transcriptomics using high-throughput data
such as microarrays and RNA-seq. We belive it would be relevant to unravel a 
robust gene expression signature for COPD patients regardless if it is from different
experiments or laboratories. 

### Background 

This script is part of an analysis performed with PulmonDB data that looks to 
determine a common deferentially expressed genes. The analysis was divided into 
different vignettes computing different steps. In this script out objective is
to do a meta-analysis with raw data downloaded from [GEO](https://www.ncbi.nlm.nih.gov/geo/). The results will be used as a reference 
and to compare them with a meta-analysis computed with PulmonDB sample contrasts.

We want to pre-process and re-analyze transcriptomic experiments from [GEO](https://www.ncbi.nlm.nih.gov/geo/) that have:

- Lung tissue samples

- COPD vs CONTROL group


## Input

This script needs the following files:

<div class="alert alert-success">
  <strong>Data 1:</strong> Matrix of normalized experiments  
  <strong>Data 2:</strong> Meta data 
</div>


### Setup

```{css, echo=FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
```

All the data has already been downloaded in the cluster *10.200.0.42*, for 
accessing:

```{bash, eval=F}
ssh ana@10.200.0.42

cd /home/ana/R/Meta-analysis_COPD
```


For running the script, type:

```{bash, eval= F}
R -e "rmarkdown::render('vignettes/RMA_DE-23-June-2020.Rmd')"
```


The script can be found in: `r getwd() `


```{r setup, class.source="bg-info"}
setwd("..")
PATH = getwd()
DATA_DIR = file.path(PATH,"data")
OUTPUT_DIR = file.path(PATH,"output_data")
FIG_DIR = file.path(PATH,"fig")
TODAY = Sys.Date()

knitr::opts_knit$set(root.dir = PATH)
knitr::opts_chunk$set(echo = TRUE)
```

And the analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(knitr) 
library(oligo)
library(tidyverse)
library(limma)
library(SummarizedExperiment)
library(GEOquery)
```

### Experiments

We selected 7 experiments that are in PulmonDB and are lung samples from COPD patients 
and that also have a control group to compare. These experiments are described 
in the following table:


```{r experiments}
gse_table <- read.csv(file.path(DATA_DIR,"GSE_table.csv"), row.names = 1)
kable(gse_table, caption = "GSE information")
```

The experiment **GSE57148** is a RNA-seq experiment, and we do not need to normalize
the data using RMA but we will download counts from [ReCount2](https://jhubiostatistics.shinyapps.io/recount/).


## Local functions

### **rawCEL_normCEL**: Pre-process raw data

We first read and pre-process raw *.CEL* files. Each experiment has it own 
folder with raw data per sample. Then we normalized using RMA algorithm and 
finally, we save it in a *.CSV* file. 

In this function you need:

<div class="alert alert-warning">
  <strong>Input:</strong> GSE ID   
  <strong>Output:</strong> Samples normalized, boxplots and histograms from raw and normalized data 
</div>



```{r rawCEL_normCEL}
rawCEL_normCEL <- function(gse){
  # select CEL files
  celfiles <- list.celfiles(file.path(DATA_DIR,"celfiles",gse), full.names=TRUE,listGzipped=TRUE)
  # read CEL files in R
  rawData <- read.celfiles(celfiles)
  #### Figures of raw data
  #pdf(str_c("raw_",gse,"_boxplot",TODAY,".pdf"))
  ## boxplot of raw data
  boxplot(rawData)
  ## hist of raw data
  hist(rawData)
  #dev.off()
  ## RMA normalization
  normData <- rma(rawData)
  #### Figures of Normalized data
  #pdf(str_c("norm_",gse,"_boxplot",TODAY,".pdf"))
  ## boxplot of norm data
  boxplot(normData)
  ## hist of norm data
  hist(normData)
  # dev.off()
  #write.csv(exprs(normData),str_c(OUTPUT_DIR,"/",gse,"_normData",TODAY,".txt"),quote=F)
  return(normData)
}

#sapply(tissue,rawCEL_normCEL)

```


### **get_GEO**: get annotation using GEOquery package

This function download the `ExpressionSet` object from `GEOquery` that has 
the sample annotations, then we replace the expression values with our calculated
pre-processed data. 

<div class="alert alert-warning">
  <strong>Input:</strong> GSE ID, *norm* object with pre-processed values   
  <strong>Output:</strong> ExpressionSet object with GEOquery annotation and 
  pre-processed values
</div>


```{r get_GEO}
get_GEO <- function(gse,norm){
  qx <- getGEO(gse)
  message("Data downloaded from GEOquery:")
  print(qx)
  qx <- qx[[1]]
  message("Colnames of GEOquery object:")
  print(colnames(qx)[1:5])
  message("Colnames of calculated pre-processed data:")
  print(colnames(norm)[1:5])
  # Rename sample columns (Change GSM18403.CEL.gz to GSM18403)
  # sort(colnames(norm1))
  colnames(norm) = colnames(qx)
  exprs(norm)[1:3,1:3]
  exprs(qx) <- exprs(norm)
  return(qx)
}

```



### **DE**: Differential expression analysis

Using this funtion, we get a table with differential expression gene results using
`limma` package for fitting a linear model to get genes differentially expressed
between a "Control" and a "COPD" group. 

<div class="alert alert-warning">
  <strong>Input:</strong> GSE ID, optional: *colCOPD* is the column name in which 
  the information of disease status can be found, *coeff* will show results of 
  contrast with coeffitient found in possiton 2   
  <strong>Output:</strong> Table of differential expression results with all
  genes 
</div>


```{r DE}
DE <- function(ExpressionSet,colCOPD="Disease",coeff= 2){
     # it creates the design matrix and performs limma
     fit <- lmFit(ExpressionSet, model.matrix(as.formula(paste("~ 1 +", colCOPD)),
                                              data = pData(ExpressionSet)))
     # eBayes in lmFit model
     ebf <- eBayes(fit)
     print(colnames(coef(fit)))
     # It gets the genes with the p-values
     res <- topTable(ebf, number = Inf, p.value = 1, coef = coeff,confint=T)
     # It formats in a tibble
     res <- as_tibble(res,rownames="rownames")
}

```


## `r rownames(gse_table)[1] `

This experiment evaluates gene expression profiles of emphysema using “usual” emphysema and Alpha-1 Antitrypsin Deficiency-related emphysema (AAD). As a control group, authors used normal lung tissue from “organs donated for transplant, but unused due to age or size mismatch”, non of those individuals were smokers or reported clinical airflow limitation.

This study showed that inflammation, immune responses, and proteolysis are emphysema characteristics. They also found similarities and differences between AAD and “usual” emphysema.


The authors measured 15 lung samples, 5 controls, 5 “usual” emphysema and 5 AAD emphysema. The raw data is in cel files, the platform used is [Hu6800] Affymetrix. The study was perfomed in Colorado, USA.

### Pre-process raw data

We pre-processed raw data using the function `rawCEL_normCEL`, plots will be 
shown as additional output.

```{r 1}
gse1<- rownames(gse_table)[1] 
norm1 <- rawCEL_normCEL(gse1)
norm1
```

### Get annotation 

We used `GEOquery` package to obtain sample annotations and our previous 
calculated pre-processed values to create an `ExpressionSet` object.

```{r}
# get annotation using GEOquery package
geo1 <- get_GEO(gse1,norm1)

```

### Select column with COPD description

Each experiment has its own annotation and we needed to look for a column 
describing which sample is a "Control" and which one is 
"COPD". 

```{r}
head(pData(geo1))
```

Names will be different but it is **important** to check that "Control" group
is the first level. If need it re-level groups.

```{r class.source="bg-danger",class.output="bg-danger" }
pData(geo1)["Disease"] <- factor(pData(geo1)[,"description"],levels = c("Normal lung","'usual' emphysemic lung","Alpha-1 Antitrypsin Deficiency-related emphysemic lung"))

table(pData(geo1)$Disease)
```

### Differential expression analysis

Using `DE()` function (described above), we performed a lineal regression model 
to calculate the logarithm fold change of all genes between a "Control" and a
"COPD" group. We also rename colnames adding the GSE ID at the end and finally, 
we save the output in a `.CSV` file. 

```{r}
de1 <- DE(geo1)

colnames(de1) <- str_c(colnames(de1),"_",gse1)
colnames(de1)

write_csv(de1,
          path=str_c(OUTPUT_DIR,"/TableGenes_",gse1,"_",TODAY,".csv")
          )
```


## `r rownames(gse_table)[2] `

Emphysema gene expression was measured using sever, mildly and non emphysema lung tissue. The authors separate samples in two groups, 18 sever and 12 mildly/non emphysema because the limited sample size (7 mildly, 5 non emphysema). From control group, 9 tissues were obtained from smokers with nodules suspicious for lung cancer.

The results of this study shows oxidative stress, extracellular matrix synthesis, and inflammation pathways overexpressed in severe emphysema, whereas endothelium-related was decreased.

The authors didn’t annotate samples individually, the meta information is not available. We assumed that N refers to controls and L to sever emphysema, unfortunately, we can’t separate mildly and non emphysema because of the lack of infomation.

Raw data is in cel files, the platform used is [HG-U133A] Affymetrix. The study was perfomed in Boston, USA (Boston University Medical Center).

### Pre-process raw data

We pre-processed raw data using the function `rawCEL_normCEL`, plots will be 
shown as additional output.

```{r 2}
gse2<- rownames(gse_table)[2] 
norm2 <- rawCEL_normCEL(gse2)
norm2
```

### Get annotation 

We used `GEOquery` package to obtain sample annotations and our previous 
calculated pre-processed values to create an `ExpressionSet` object.

```{r}
# get annotation using GEOquery package
geo2 <- get_GEO(gse2,norm2)

```

### Select column with COPD description

Each experiment has its own annotation and we need to look for a column 
describing which sample is a "Control" and which one is 
"COPD". 

```{r}
head(pData(geo2))
```

Names will be different but it is **important** to check that "Control" group
is the first level. If need it re-level groups.

```{r class.source="bg-danger",class.output="bg-danger" }
pData(geo2)["Disease"] <- str_count(as.character(pData(geo2)[,"title"]),"L")
table(pData(geo2)$Disease)

```

### Differential expression analysis

Using `DE()` function (described above), we performed a lineal regression model 
to calculate the logarithm fold change of all genes between a "Control" and a
"COPD" group. We also rename colnames adding the GSE ID at the end and finally, 
we save the output in a `.CSV` file. 

```{r}
de2 <- DE(geo2)

colnames(de2) <- str_c(colnames(de2),"_",gse2)
colnames(de2)

write_csv(de2,
          path=str_c(OUTPUT_DIR,"/TableGenes_",gse2,"_",TODAY,".csv")
          )
```





## `r rownames(gse_table)[3] `

This experiment measured 8 lung tissue samples from 8 different regions, in total they had 64 gene expression samples, 6 sever COPD patients and 2 donors. The control samples were lung donation organs without suitable recipient, one was a smoker and the other one never smoked.

Results showed inflammation over expressed and tissue repair under expressed in emphysema.

In this GSE id, we can find two experiments that uses different platforms, one measured COPD patients and controls using [HuEx-1_0-st] Affymetrix. And the other one measured fibroblast cell line with [HuGene10stv1_Hs_ENSG] Affymetrix. Authors used Human lung fibroblast cultures (HFL-1) with two concentrations of GHK or with TGFβ1.


### Pre-process raw data

We pre-processed raw data using the function `rawCEL_normCEL`, plots will be 
shown as additional output.

```{r 3}
gse3<- rownames(gse_table)[3] 
norm3 <- rawCEL_normCEL(gse3)
norm3
```

### Get annotation 

We used `GEOquery` package to obtain sample annotations and our previous 
calculated pre-processed values to create an `ExpressionSet` object.

```{r}
# get annotation using GEOquery package
geo3 <- get_GEO(gse3,norm3)

```

### Select column with COPD description

Each experiment has its own annotation and we needed to look for a column 
describing which sample is a "Control" and which one is 
"COPD". 

```{r}
head(pData(geo3))
```

Names will be different but it is **important** to check that "Control" group
is the first level. If need it re-level groups.

```{r class.source="bg-danger",class.output="bg-danger" }
pData(geo3)["Disease"] <- pData(geo3)[,"characteristics_ch1.3"]

table(pData(geo3)$Disease)
```

### Differential expression analysis

Using `DE()` function (described above), we performed a lineal regression model 
to calculate the logarithm fold change of all genes between a "Control" and a
"COPD" group. We also rename colnames adding the GSE ID at the end and finally, 
we save the output in a `.CSV` file. 

```{r}
de3 <- DE(geo3)

colnames(de3) <- str_c(colnames(de3),"_",gse3)
colnames(de3)

write_csv(de3,
          path=str_c(OUTPUT_DIR,"/TableGenes_",gse3,"_",TODAY,".csv")
          )
```



## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Tables with DE results:</strong> Tables with log fold change and p-values 
  calculated   
  <strong>Table of merged results:</strong> Table with all DE results 
</div>

## Session Info

```{r sessionInfo, attr.output='style="max-height: 500px;"'}
sessionInfo()
```





