---
title: "Enrichment Analysis"
author: Ana BVA
date: "`r BiocStyle::doc_date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

## Introduction

<div class="alert alert-info">
  <strong>Aim:</strong> The aim is to do an enrichment analysis using results from  
  the meta-analysis.
</div>

### Background 

This analysis is based on [hyperR BioC 2020 workshop](https://montilab.github.io/hypeR-workshop/articles/hyper_workshop.html).

## Input

This script needs the following files:

<div class="alert alert-warning">
  <strong>Data 1:</strong> Matrix of normalized experiments  
  <strong>Data 2:</strong> Meta data 
</div>

## Analysis

### Setup

The script can be found in: `r getwd() `

```{r setup}
PATH = here::here() # "."
DATA_DIR = file.path(PATH,"Data/")
OUTPUT_DIR = file.path(PATH,"output/")
FIG_DIR = file.path(PATH,"Fig")

knitr::opts_knit$set(root.dir = PATH)
```

And this analysis is run in: `r getwd() `

### Libraries

```{r libraries, message=FALSE, warning=FALSE}
# BiocManager::install("hypeR")
library(hypeR)
library(WGCNA)
library(tidyverse)
library(reactable)
```


### STEP 1

```{r , class.source="bg-info",class.output="bg-info" }
meta <- read.csv("output_data/meta-analysis_preprocessData-2020-08-13.csv")[,-1]

signature <- meta[meta$qval <= 0.1,"QEp"]
names(signature) <- meta[meta$qval <= 0.1,"genes"]

head(signature)
```

### STEP 2

We are importing the latest genesets from Kegg.

```{r}
# KEGG
genesets <- hyperdb_rgsets("REACTOME", version="70.0")
genesets <- hyperdb_rgsets("KEGG", "92.0")

#HALLMARL
HALLMARK <- msigdb_gsets(species="Homo sapiens", category="H",clean = T)
print(HALLMARK)

#ENRICHR
available <- enrichr_available()
reactable(available)
enrichR <- enrichr_gsets("TRANSFAC_and_JASPAR_PWMs")
enrichR <- enrichr_gsets("Transcription_Factor_PPIs")
enrichR<- enrichr_gsets("GeneSigDB")
enrichR<- enrichr_gsets(available[67,1])
enrichR<- enrichr_gsets(available[68,1])
enrichR<- enrichr_gsets(available[69,1])
enrichR<- enrichr_gsets(available[71,1])
enrichR<- enrichr_gsets(available[103,1])
enrichR<- enrichr_gsets(available[79,1])
enrichR<- enrichr_gsets(available[80,1])
enrichR<- enrichr_gsets(available[102,1])
enrichR<- enrichr_gsets(available[107,1])

enrichR<- enrichr_gsets(available[1,1])
enrichR<- enrichr_gsets(available[130,1])
enrichR<- enrichr_gsets(available[110,1])
enrichR<- enrichr_gsets(available[132,1])
```


```{r}
hyp_obj <- hypeR(names(signature),KEGG, test="hypergeometric", background=nrow(meta), fdr=0.5, plotting=TRUE)
hyp_obj$plots[[1]]
print(hyp_obj)
```

```{r}
hyp_obj <- hypeR(signature, genesets, test="kstest", fdr=0.5, plotting=TRUE)
hyp_obj$plots[[1]]
print(hyp_obj)
```

```{r}
# Show interactive table
hyp_show(hyp_obj)

# Plot dots plot
hyp_dots(hyp_obj) + theme_minimal()

# Plot enrichment map
hyp_emap(hyp_obj)

# Plot hiearchy map
hyp_hmap(hyp_obj)

# Save to excel
hyp_to_excel(hyp_obj)

# Save to table
hyp_to_table(hyp_obj)

# Generate markdown report
hyp_to_rmd(hyp_obj,"hypr-KEGG-genesets.rmd")
hyp_to_rmd(hyp_obj,"hypr-TRANSFAC_and_JASPAR_PWMs.rmd")
```


Enrichment analysis following this tutorial https://yulab-smu.github.io/clusterProfiler-book/chapter12.html#dot-plot

```{r}
# TF enrichment and TF gene expression 
#enr <- gsub(" .*","",hyp_obj$as.data.frame()$label)
#topgenes[which(rownames(topgenes) %in% enr),]

library(DOSE)
library(org.Hs.eg.db)
library(enrichplot)


signature <- meta[meta$qval <= 0.1,"QEp"]
names(signature) <- meta[meta$qval <= 0.1,"genes"]

eid <- select(org.Hs.eg.db, names(signature), "ENTREZID", "SYMBOL")[["ENTREZID"]]

foldch <- full_tables[,c("Gene.Symbol",FC)]
genefc <- rowMeans(foldch[foldch$Gene.Symbol %in% names(signature),-1])
names(genefc) <- names(signature)
names(genefc) <- eid
genefc <- na.omit(genefc)

library(pathview)
library(clusterProfiler)

kk <- enrichKEGG(gene         = names(genefc),
                 organism     = 'hsa',
                 pvalueCutoff = 0.05)
head(kk)
barplot(kk)

kk2 <- gseKEGG(geneList     = sort(na.omit(genefc),decreasing = T),
               organism     = 'hsa',
               nPerm        = 1000,
               minGSSize    = 10,
               pvalueCutoff = 0.05,
               verbose      = FALSE)
head(kk2)
dotplot(kk2) + ggtitle("dotplot for GSEA")
ridgeplot(kk2)
gseaplot2(kk2, geneSetID = 1, title = kk2$Description[1],pvalue_table = TRUE)

pathview(gene.data  = genefc*2,
         pathway.id = "hsa04151",
         species    = "hsa",
         limit      = list(gene=max(abs(genefc)), cpd=1))

```


## Output

This script produces the following data, and can be found in `r getwd() ` 

<div class="alert alert-success">
  <strong>Figure 1:</strong> Boxplot of samples  
  <strong>Figure 2:</strong> Heatmap of COPD patients 
</div>

## Session Info

```{r sessionInfo}
sessionInfo()
```





